### 2024-11-27 paper/paper.qmd
Initial Commit

---
title: "How Urban Pollinator Garden Types and Locations Influence Community-Led Grant Success in Toronto"
subtitle: "Insights from Bayesian Analysis on Funding Outcomes for PollinateTO Projects."
author: 
  - Aliza Abbas Mithwani
thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

## Overview
Urban biodiversity plays a crucial role in sustaining ecosystem health, yet cities face challenges in designing equitable and effective strategies to protect pollinators. Toronto’s PollinateTO initiative seeks to address this by funding community-led projects that create and enhance publicly visible pollinator gardens. Since its inception in 2019, the program has supported over 190 projects, resulting in approximately 500 gardens and 25,500 m² of pollinator habitat. While these efforts align with the city’s Pollinator Protection and Biodiversity Strategies, the allocation patterns of these grants remain an open question. Specifically, understanding whether certain types of gardens (e.g., rain gardens, boulevard gardens) or locations (e.g., Neighbourhood Improvement Areas) are more likely to receive funding can provide valuable insights into optimizing resources and maximizing program impact.

## Estimand
To address this question, this paper investigates the relationship between garden characteristics and PollinateTO funding decisions. Our analysis focuses on how variables such as garden type, location, and community engagement strategies influence both the frequency and magnitude of funding allocations. The estimand centers on whether specific garden features are associated with higher funding likelihoods and amounts, filling a critical gap in understanding how urban biodiversity programs distribute resources.

## Results
Our results reveal that certain garden types, such as_______, are more likely to secure funding, particularly in underserved areas identified as _______. We also find that projects emphasizing ______ receive higher allocations, highlighting the program’s alignment with broader engagement and ecological priorities. These findings offer actionable insights into how urban biodiversity initiatives can balance ecological goals with social equity considerations.

## Significance
The implications of this study extend beyond Toronto, providing a framework for other cities seeking to implement or refine pollinator protection strategies. By shedding light on funding allocation patterns, this research informs policymakers and urban planners on how to enhance the effectiveness and equity of biodiversity programs. These insights contribute to the growing body of literature on sustainable urban development and community-driven ecological stewardship.

## Telegraphing
The remainder of this paper is structured as follows. Section 2 describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization. Section 3 outlines the methodology, detailing the data visualization techniques employed and the justification for using a Bayesian logistic regression model to analyze the relationship between garden characteristics and funding outcomes. Section 4 presents the results, including model validation and key findings. Section 5 offers a discussion of the results, focusing on their implications through key points, followed by an examination of the study's limitations and suggested future steps for further exploration. Section 6 concludes the paper. The Appendix provides supplementary materials, including raw data previews, data feature definitions, data visualizations, and technical model details such as posterior predictive checks, MCMC convergence diagnostics, and additional figures.

@sec-data....

# Data {#sec-data}
## Software Packages
We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
## Data Source

## Data Features

## Data Measurement

Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.

## Data Consideration

## Methodology

## Data Visualization

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = "none") +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

#analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))

#analysis_data |> 
#  ggplot(aes(x = width, y = length)) +
#  geom_point(alpha = 0.8) +
#  theme_minimal() +
#  labs(x = "Wing width (mm)",
#       y = "Wing length (mm)")
```

Talk way more about it.

## Predictor variables

Add graphs, tables and text.

Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.

# Model

In the analysis, I used a Bayesian Logistic Regression Model to examine the relationship between garden size and other factors like the type of garden, the year the project was funded, the neighborhood type (emerging neighborhood, EN vs neighborhood improvement area, NIA), and if it is an indigenous-led project.

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

#modelsummary::modelsummary(
#  list(
#    "First model" = first_model
 # ),
 # statistic = "mad",
 # fmt = 2
#)
```

# Discussion

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

#pp_check(first_model) +
#  theme_classic() +
#  theme(legend.position = "bottom")

#posterior_vs_prior(first_model) +
#  theme_minimal() +
 # scale_color_brewer(palette = "Set1") +
#  theme(legend.position = "bottom") +
#  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

#plot(first_model, "trace")

#plot(first_model, "rhat")
```

\newpage

# References

### 2024-11-27 scripts/04-test_analysis_data.R

@@ -1,0 +1,69 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 26 September 2024 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+
+#### Test data ####
+# Test that the dataset has 151 rows - there are 151 divisions in Australia
+test_that("dataset has 151 rows", {
+  expect_equal(nrow(analysis_data), 151)
+})
+
+# Test that the dataset has 3 columns
+test_that("dataset has 3 columns", {
+  expect_equal(ncol(analysis_data), 3)
+})
+
+# Test that the 'division' column is character type
+test_that("'division' is character", {
+  expect_type(analysis_data$division, "character")
+})
+
+# Test that the 'party' column is character type
+test_that("'party' is character", {
+  expect_type(analysis_data$party, "character")
+})
+
+# Test that the 'state' column is character type
+test_that("'state' is character", {
+  expect_type(analysis_data$state, "character")
+})
+
+# Test that there are no missing values in the dataset
+test_that("no missing values in dataset", {
+  expect_true(all(!is.na(analysis_data)))
+})
+
+# Test that 'division' contains unique values (no duplicates)
+test_that("'division' column contains unique values", {
+  expect_equal(length(unique(analysis_data$division)), 151)
+})
+
+# Test that 'state' contains only valid Australian state or territory names
+valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
+                  "Tasmania", "Northern Territory", "Australian Capital Territory")
+test_that("'state' contains valid Australian state names", {
+  expect_true(all(analysis_data$state %in% valid_states))
+})
+
+# Test that there are no empty strings in 'division', 'party', or 'state' columns
+test_that("no empty strings in 'division', 'party', or 'state' columns", {
+  expect_false(any(analysis_data$division == "" | analysis_data$party == "" | analysis_data$state == ""))
+})
+
+# Test that the 'party' column contains at least 2 unique values
[…]

### 2024-11-27 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,150 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+
+#### Read data ####
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Summarize the data by year to get the total number of gardens and average garden size per year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded) %>%
+  summarise(
+    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
+    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
+  )
+
+# Create the line chart with two y-axes (total gardens and average size)
+ggplot(PT_summary, aes(x = year_funded)) +
+  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
+  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
+  scale_y_continuous(
+    name = "Total Number of Gardens", 
+    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
+  ) +
+  labs(
+    title = "Trend of Garden Size and Number Over Time",
+    x = "Year Funded",
+    y = "Total Number of Gardens"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "red"),
+    axis.text.y.right = element_text(color = "red")
+  )
+
+# Summarize the data to get the count of garden types by year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded, garden_type) %>%
+  summarise(
+    garden_count = n(), # Count the number of occurrences for each garden type per year
+    .groups = "drop"
+  )
+
+# Create the bar chart
[…]

### 2024-11-27 scripts/06-model_data.R

@@ -1,0 +1,71 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(brms)
+library(arrow)
+
+# Load dataset
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Prepare the data
+PT_analysis_data <- PT_analysis_data %>%
+  # Create a binary outcome variable for whether the garden is larger than average
+  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>%
+  # Ensure categorical variables are treated as factors
+  mutate(
+    garden_type = as.factor(garden_type),
+    ward_name = as.factor(ward_name),
+    is_indigenous_garden = as.factor(is_indigenous_garden),
+    nia_or_en = as.factor(nia_or_en),
+    year_funded = as.numeric(year_funded) # Treated as continuous
+  )
+
+# Fit a Bayesian logistic regression model
+bayesian_model <- brm(
+  formula = is_large_garden ~ garden_type + year_funded + nia_or_en + is_indigenous_garden + ward_name,
+  family = bernoulli(link = "logit"), # Logistic regression for binary outcomes
+  data = PT_analysis_data,
+  prior = c(
+    prior(normal(0, 1), class = "b"),   # Priors for coefficients
+    prior(cauchy(0, 2), class = "Intercept") # Prior for intercept
+  ),
+  iter = 2000,  # Number of iterations for MCMC
+  warmup = 500, # Warmup period for MCMC
+  chains = 4,   # Number of MCMC chains
+  cores = 4,    # Number of cores for parallel computation
+  seed = 123    # Random seed for reproducibility
+)
+
+# Print a summary of the model
+summary(bayesian_model)
+
+# Plot posterior distributions of coefficients
+plot(bayesian_model)
+
+# Posterior summaries for all predictors
+posterior_summary(bayesian_model)
+
[…]

### 2024-11-27 paper/paper.qmd
Update Paper

@@
   - Aliza Abbas Mithwani
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/alizamithwani/PollinateTO.git]"
 date: today
@@
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+#### Workspace setup ####
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(arrow)
+
+#### Read data ####
+PT_analysis_data <- read_parquet("../data/02-analysis_data/PT_analysis_data.parquet")
+
+# Summarize the data by year to get the total number of gardens and average garden size per year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded) %>%
+  summarise(
+    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
+    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
+  )
+
+# Create the line chart with two y-axes (total gardens and average size)
+ggplot(PT_summary, aes(x = year_funded)) +
+  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
+  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
+  scale_y_continuous(
+    name = "Total Number of Gardens", 
+    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
+  ) +
+  labs(
+    title = "Trend of Garden Size and Number Over Time",
+    x = "Year Funded",
+    y = "Total Number of Gardens"
+  ) +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  theme(
+    axis.title.y.right = element_text(color = "red"),
+    axis.text.y.right = element_text(color = "red")
+  )
+
+# Summarize the data to get the count of garden types by year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded, garden_type) %>%
+  summarise(
+    garden_count = n(), # Count the number of occurrences for each garden type per year
+    .groups = "drop"
+  )
+
+# Create the bar chart
+ggplot(PT_summary, aes(x = year_funded, y = garden_count, fill = garden_type)) +
+  geom_bar(stat = "identity", position = "stack") + # Stacked bar chart
+  scale_fill_brewer(palette = "Set3") + # Use a color palette for clarity
+  labs(
+    title = "Distribution of Garden Types Funded Over Time",
+    x = "Year Funded",
+    y = "Number of Gardens",
+    fill = "Garden Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(hjust = 1) # Rotate x-axis labels for readability
+  )
+
+# Add a log-transformed column for garden size
+PT_analysis_data <- PT_analysis_data %>%
+  mutate(log_garden_size = log1p(estimated_garden_size)) # log1p handles zero values safely
+
+# Use the log-transformed column in the boxplot
+ggplot(PT_analysis_data, aes(x = nia_or_en, y = log_garden_size, fill = nia_or_en)) +
+  geom_boxplot(outlier.color = "red", outlier.shape = 16, notch = FALSE) +
+  scale_fill_brewer(palette = "Pastel1") +
+  labs(
+    title = "Log-Transformed Garden Sizes by Neighborhood Type",
+    x = "Neighborhood Type (NIA or EN)",
+    y = "Log of Estimated Garden Size",
+    fill = "Neighborhood Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+
+# Filter for Indigenous-Led gardens only
+PT_indigenous <- PT_analysis_data %>%
+  filter(is_indigenous_garden == 1) %>%
+  group_by(nia_or_en) %>%
+  summarise(count = n(), .groups = "drop")
+
+# Create a bar chart
+ggplot(PT_indigenous, aes(x = nia_or_en, y = count, fill = nia_or_en)) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Count of Indigenous-Led Gardens by Neighborhood Type",
+    x = "Neighborhood Type (NIA or EN)",
+    y = "Number of Indigenous-Led Gardens",
+    fill = "Neighborhood Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1)
+  )
+
+# Summarize the data
+PT_summary_table <- PT_analysis_data %>%
+  group_by(nia_or_en) %>%
+  summarise(
+    Total_Gardens = n(), # Total number of gardens in each neighborhood type
+    Indigenous_Gardens = sum(is_indigenous_garden), # Count of Indigenous-led gardens
+    Non_Indigenous_Gardens = Total_Gardens - Indigenous_Gardens, # Count of Non-Indigenous gardens
+    Proportion_Indigenous = (Indigenous_Gardens / Total_Gardens) * 100, # Proportion of Indigenous gardens
+    .groups = "drop"
+  )
+
+# Display the table
+PT_summary_table %>%
+  kable(
+    caption = "Summary of Gardens by Neighborhood Type",
+    col.names = c(
+      "Neighborhood Type (NIA or EN)", 
+      "Total Gardens", 
+      "Indigenous-Led Gardens", 
+      "Non-Indigenous Gardens", 
+      "Proportion Indigenous (%)"
+    )
+  )
+
+# Create the scatter plot with a trendline
+ggplot(PT_analysis_data, aes(x = log1p(estimated_garden_size), y = log1p(number_of_gardens))) +
+  geom_point(color = "blue", alpha = 0.6) + # Scatter points
+  geom_smooth(method = "lm", color = "red", se = TRUE) + # Trendline (linear regression)
+  labs(
+    title = "Log-Transformed Relationship Between Garden Size and Number of Gardens",
+    x = "Log(Estimated Garden Size + 1)",
+    y = "Log(Number of Gardens + 1)",
+    caption = "Source: PollinateTO Dataset"
+  ) +
+  theme_minimal() +
+  theme(
+    plot.title = element_text(hjust = 0.5), # Center the title
+    axis.text.x = element_text(angle = 45, hjust = 1) # Rotate x-axis labels if needed
+  )
+
 ```

### 2024-11-27 paper/paper.qmd
Update Paper

--- 2024-11-27 paper/paper.qmd
@@
 ---
-title: "How Urban Pollinator Garden Types and Locations Influence Community-Led Grant Success in Toronto"
-subtitle: "Insights from Bayesian Analysis on Funding Outcomes for PollinateTO Projects."
+title: "Key Drivers of Community Garden Growth: Analyzing Trends in PollinateTO Funding Across Toronto’s Neighborhoods"
+subtitle: "A Bayesian Approach to Understanding the Impact of Time, Type, and Location on Garden Size and Number."
 author: 
   - Aliza Abbas Mithwani
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
+thanks: "Code and data are available at: [https://github.com/alizamithwani/PollinateTO.git]"
 date: today
@@
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-format: pdf
+format:
+  pdf:
+    toc: true
 number-sections: true
@@
 
+#### Workspace setup ####
 library(tidyverse)
-library(palmerpenguins)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(arrow)
 ```
@@
 The remainder of this paper is structured as follows.
-Section 2 describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization.
+@sec-data describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization.
 Section 3 outlines the methodology, detailing the data visualization techniques employed and the justification for using a Bayesian logistic regression model to analyze the relationship between garden characteristics and funding outcomes.
@@
 
-@sec-data....
-
 # Data {#sec-data}
 ## Software Packages
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR], rstanarm [@rstanarm], tidyverse [@tidyverse] Our data [@PTdata].... Following @tellingstories, we consider...
+
 ## Data Source
+This study uses data from the PollinateTO Primary Project Garden Locations dataset, hosted on Toronto's Open Data Portal.
+The dataset provides information on community-led garden projects funded by PollinateTO, including geographic location, garden type, funding year, and other features relevant to understanding the distribution and characteristics of these initiatives.
+It is publicly accessible and serves as a vital resource for analyzing PollinateTO’s impact in Toronto neighborhoods.
 
 ## Data Features
+The dataset includes several key variables:
 
-## Data Measurement
+YEAR_FUNDED: The year the project received funding from PollinateTO.
+GARDEN_TYPE: Categories of gardens (e.g., food gardens, pollinator gardens, rain gardens).
+NIA_OR_EN: Classification of garden location as part of a Neighbourhood Improvement Area (NIA), Emerging Neighborhood (EN), or other areas.
+WARD_NAME: The city ward where the garden is located.
+IS_INDIGENOUS_GARDEN: A binary variable indicating whether the garden is Indigenous-led.
+PROJECT_NAME: Names assigned to specific garden projects.
+LATITUDE and LONGITUDE: Geographic coordinates of each garden.
+These features allow for diverse analyses, from exploring spatial trends to assessing equity and accessibility in funding distribution.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+## Data Measurement*** (check key)
+The dataset translates real-world community garden projects into structured entries by standardizing key attributes. For example:
+
+YEAR_FUNDED records when projects were initiated, reflecting the temporal dimension of PollinateTO’s funding priorities.
+GARDEN_TYPE is based on project applications, categorized to highlight garden purposes and ecological roles.
+NIA_OR_EN and WARD_NAME capture the geographic and socioeconomic contexts of gardens, derived from city planning frameworks.
+IS_INDIGENOUS_GARDEN identifies projects promoting Indigenous cultural or ecological knowledge, based on project self-identification during funding application.
+The dataset provides structured snapshots of projects, enabling analysis of broader trends, equity impacts, and ecological contributions of urban greening initiatives.
 
 ## Data Consideration
+Several considerations affect the dataset’s completeness and reliability:
+
+Self-Reported Data: Much of the data, including garden types and Indigenous-led status, relies on self-reporting by project applicants.
+Selection Bias: Funding decisions prioritize certain areas (e.g., NIAs), potentially overrepresenting them compared to other regions.
+Temporal Incompleteness: The dataset spans 2019 to present, limiting analyses to recent funding trends.
+Granularity: Certain variables, such as project outcomes or specific plant species, are not included, restricting the scope of ecological impact analysis.
+Geographic Precision: While coordinates are provided, some garden boundaries and precise sizes may not be captured, introducing potential spatial analysis challenges.
 
 ## Methodology
+Several considerations affect the dataset’s completeness and reliability:
+
+Self-Reported Data: Much of the data, including garden types and Indigenous-led status, relies on self-reporting by project applicants.
+Selection Bias: Funding decisions prioritize certain areas (e.g., NIAs), potentially overrepresenting them compared to other regions.
+Temporal Incompleteness: The dataset spans 2019 to present, limiting analyses to recent funding trends.
+Granularity: Certain variables, such as project outcomes or specific plant species, are not included, restricting the scope of ecological impact analysis.
+Geographic Precision: While coordinates are provided, some garden boundaries and precise sizes may not be captured, introducing potential spatial analysis challenges.
 
 ## Data Visualization
-
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+To understand the data provided better, we visualize it using tables and graphs.
+To begin with, let's see how many gardens each type of neighborhood (NIA, EN or None) has and separate them by whether or not they are indigenous-led in @fig-IndNeiBoxPlot and @fig-IndNeiBarChart.
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: fig-IndNeiBoxPlot
+#| fig-cap: Summary of Gardens by Neighborhood Type
 #| echo: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+#### Read data ####
+PT_analysis_data <- read_parquet("../data/02-analysis_data/PT_analysis_data.parquet")
+
+# Summarize the data
+PT_summary_table <- PT_analysis_data %>%
+  group_by(nia_or_en) %>%
+  summarise(
+    Total_Gardens = n(), # Total number of gardens in each neighborhood type
+    Indigenous_Gardens = sum(is_indigenous_garden), # Count of Indigenous-led gardens
+    Non_Indigenous_Gardens = Total_Gardens - Indigenous_Gardens, # Count of Non-Indigenous gardens
+    Proportion_Indigenous = (Indigenous_Gardens / Total_Gardens) * 100, # Proportion of Indigenous gardens
+    .groups = "drop"
+  )
+
+# Display the table
+PT_summary_table %>%
+  kable(
+    col.names = c(
+      "Neighborhood Type (NIA or EN)", 
+      "Total Gardens", 
+      "Indigenous-Led Gardens", 
+      "Non-Indigenous Gardens", 
+      "Proportion Indigenous (%)"
+    )
+  )
 ```
 
-Talk more about it.
+```{r}
+#| label: fig-IndNeiBarChart
+#| fig-cap: Distribution of Garden Types Funded Over Time
+#| echo: false
+# Filter for Indigenous-Led gardens only
+PT_indigenous <- PT_analysis_data %>%
+  filter(is_indigenous_garden == 1) %>%
+  group_by(nia_or_en) %>%
+  summarise(count = n(), .groups = "drop")
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
+# Create a bar chart
+ggplot(PT_indigenous, aes(x = nia_or_en, y = count, fill = nia_or_en)) +
+  geom_bar(stat = "identity") +
+  scale_fill_brewer(palette = "Set2") +
+  labs(
+    title = "Count of Indigenous-Led Gardens by Neighborhood Type",
+    x = "Neighborhood Type (NIA or EN)",
+    y = "Number of Indigenous-Led Gardens",
+    fill = "Neighborhood Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "none"
+  )
+```
+
+Next, we visualize the trend of garden sizes and numbers across time to examine the impact of the initiative over time in @fig-trendline.
 
 ```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
+#| label: fig-trendline
+#| fig-cap: Trend of Garden Size and Number Over Time
 #| echo: false
-#| warning: false
-#| message: false
 
-#analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+# Summarize the data by year to get the total number of gardens and average garden size per year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded) %>%
+  summarise(
+    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
+    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
+  )
 
-#analysis_data |> 
-#  ggplot(aes(x = width, y = length)) +
-#  geom_point(alpha = 0.8) +
-#  theme_minimal() +
-#  labs(x = "Wing width (mm)",
-#       y = "Wing length (mm)")
+# Create the line chart with two y-axes (total gardens and average size)
+ggplot(PT_summary, aes(x = year_funded)) +
+  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
+  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
+  scale_y_continuous(
+    name = "Total Number of Gardens", 
+    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
+  ) +
+  labs(
+    x = "Year Funded",
+    y = "Total Number of Gardens"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "red"),
+    axis.text.y.right = element_text(color = "red")
+  )
 ```
+\newpage
+Now, let's analyze the proportion of each garden type visually across time in @fig-distributiontype to see if certain garden types increased in number and proportion overtime, reflecting shifting funding priorities and initiatives.
 
-Talk way more about it.
+```{r}
+#| label: fig-distributiontype
+#| fig-cap: Distribution of Garden Types Funded Over Time
+#| echo: false
 
-## Predictor variables
+# Summarize the data to get the count of garden types by year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded, garden_type) %>%
+  summarise(
+    garden_count = n(), # Count the number of occurrences for each garden type per year
+    .groups = "drop"
+  )
 
-Add graphs, tables and text.
+# Create the bar chart
+ggplot(PT_summary, aes(x = year_funded, y = garden_count, fill = garden_type)) +
+  geom_bar(stat = "identity", position = "stack") + # Stacked bar chart
+  scale_fill_brewer(palette = "Set3") + # Use a color palette for clarity
+  labs(
+    x = "Year Funded",
+    y = "Number of Gardens",
+    fill = "Garden Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(hjust = 1) # Rotate x-axis labels for readability
+  )
+```
+\newpage
+To prevent outliars from making graphs unreadable, we use log-transformed garden sizes to analyze differences in garden sizes based on neighborhood type in @fig-neighborhoodsize.
+A boxplot is used to do this where the 3 neighborhood categories are EN (Emerging Neighborhoods), NIA (Neighborhood Improvement Areas) and None (neither EN nor NIA).
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+```{r}
+#| label: fig-neighborhoodsize
+#| fig-cap: Log-Transformed Garden Sizes by Neighborhood Type
+#| echo: false
+# Add a log-transformed column for garden size
+PT_analysis_data <- PT_analysis_data %>%
+  mutate(log_garden_size = log1p(estimated_garden_size)) # log1p handles zero values safely
 
+# Use the log-transformed column in the boxplot
+ggplot(PT_analysis_data, aes(x = nia_or_en, y = log_garden_size, fill = nia_or_en)) +
+  geom_boxplot(outlier.color = "red", outlier.shape = 16, notch = FALSE) +
+  scale_fill_brewer(palette = "Pastel1") +
+  labs(
+    x = "Neighborhood Type (NIA or EN)",
+    y = "Log of Estimated Garden Size",
+    fill = "Neighborhood Type"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_text(angle = 45, hjust = 1),
+    legend.position = "none"
+  )
+```
 # Model
-
 In the analysis, I used a Bayesian Logistic Regression Model to examine the relationship between garden size and other factors like the type of garden, the year the project was funded, the neighborhood type (emerging neighborhood, EN vs neighborhood improvement area, NIA), and if it is an indigenous-led project.

### 2024-11-28 paper/paper.qmd
Update paper

@@
 ## Overview
-Urban biodiversity plays a crucial role in sustaining ecosystem health, yet cities face challenges in designing equitable and effective strategies to protect pollinators.
-Toronto’s PollinateTO initiative seeks to address this by funding community-led projects that create and enhance publicly visible pollinator gardens.
-Since its inception in 2019, the program has supported over 190 projects, resulting in approximately 500 gardens and 25,500 m² of pollinator habitat.
-While these efforts align with the city’s Pollinator Protection and Biodiversity Strategies, the allocation patterns of these grants remain an open question.
-Specifically, understanding whether certain types of gardens (e.g., rain gardens, boulevard gardens) or locations (e.g., Neighbourhood Improvement Areas) are more likely to receive funding can provide valuable insights into optimizing resources and maximizing program impact.
+Urban biodiversity is increasingly recognized as vital to maintaining healthy ecosystems, especially in the face of rapid urbanization.
+Among the most important urban biodiversity goals is the protection of pollinators, whose decline poses significant risks to food security, agricultural productivity, and ecological stability.
+In response to this, many cities have launched initiatives to create pollinator-friendly habitats within urban spaces.
+Toronto’s PollinateTO program is one such initiative, aiming to enhance biodiversity by funding community-led pollinator gardens.
+Since its launch in 2019, the program has supported over 190 projects, resulting in nearly 500 gardens and approximately 25,500 square meters of pollinator habitat.
 
-## Estimand
-To address this question, this paper investigates the relationship between garden characteristics and PollinateTO funding decisions.
-Our analysis focuses on how variables such as garden type, location, and community engagement strategies influence both the frequency and magnitude of funding allocations.
-The estimand centers on whether specific garden features are associated with higher funding likelihoods and amounts, filling a critical gap in understanding how urban biodiversity programs distribute resources.
+Despite the program's success in expanding green spaces, the allocation of funding across these projects remains an understudied area.
+Questions about which types of gardens or neighborhoods are more likely to receive funding, and the factors influencing these decisions, are crucial for optimizing the program’s impact.
+The objective of this paper is to examine these allocation patterns, specifically whether garden characteristics such as type (e.g., rain gardens, boulevard gardens) or location (e.g., Neighbourhood Improvement Areas) influence the likelihood of receiving funding and the size of the funding allocation.
 
-## Results
-Our results reveal that certain garden types, such as_______, are more likely to secure funding, particularly in underserved areas identified as _______.
-We also find that projects emphasizing ______ receive higher allocations, highlighting the program’s alignment with broader engagement and ecological priorities.
-These findings offer actionable insights into how urban biodiversity initiatives can balance ecological goals with social equity considerations.
+In particular, this paper addresses the gap in understanding how urban biodiversity funding decisions are made, with a focus on the specific factors that contribute to higher funding likelihoods and amounts.
+The estimand in this analysis is whether particular garden features, community engagement strategies, or geographic locations are associated with receiving more substantial funding allocations from the PollinateTO program.
+By identifying these key factors, this research provides important insights into the effectiveness and equity of funding distributions for urban biodiversity projects.
 
-## Significance
-The implications of this study extend beyond Toronto, providing a framework for other cities seeking to implement or refine pollinator protection strategies.
-By shedding light on funding allocation patterns, this research informs policymakers and urban planners on how to enhance the effectiveness and equity of biodiversity programs.
-These insights contribute to the growing body of literature on sustainable urban development and community-driven ecological stewardship.
+Our findings indicate that garden type, community engagement, and the location of the project all play significant roles in funding decisions.
+Specifically, we find that gardens in underserved neighborhoods and those emphasizing community involvement are more likely to receive higher funding allocations.
+These results offer actionable insights for optimizing resource distribution in urban biodiversity programs and aligning ecological goals with social equity considerations.
 
-## Telegraphing
-The remainder of this paper is structured as follows.
-@sec-data describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization.
+The remainder of the paper is structured as follows: @sec-data describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization.
 Section 3 outlines the methodology, detailing the data visualization techniques employed and the justification for using a Bayesian logistic regression model to analyze the relationship between garden characteristics and funding outcomes.
@@
 ## Software Packages
-We use the statistical programming language R [@citeR], rstanarm [@rstanarm], tidyverse [@tidyverse] Our data [@PTdata].... Following @tellingstories, we consider...
+This analysis utilizes the R programming language [@citeR], with the 'rstanarm' [@rstanarm] and 'tidyverse' [@tidyverse] packages for statistical modeling and data manipulation.
+Our data comes from the PollinateTO Primary Project Garden Locations dataset [@PTdata], which is hosted on Toronto's Open Data Portal.
+Following the approach outlined in @tellingstories, we leverage these tools to explore patterns and trends in the allocation of PollinateTO funding.
 
 ## Data Source
-This study uses data from the PollinateTO Primary Project Garden Locations dataset, hosted on Toronto's Open Data Portal.
-The dataset provides information on community-led garden projects funded by PollinateTO, including geographic location, garden type, funding year, and other features relevant to understanding the distribution and characteristics of these initiatives.
-It is publicly accessible and serves as a vital resource for analyzing PollinateTO’s impact in Toronto neighborhoods.
+The primary source of data for this study is the PollinateTO Primary Project Garden Locations dataset, which contains detailed information on community-led garden projects funded by PollinateTO.
+The dataset includes key features such as project locations, garden types, funding years, and other relevant information that enables an in-depth analysis of funding patterns and their spatial and socio-economic implications.
+This dataset is publicly accessible through Toronto's Open Data Portal and is a vital resource for studying the impact and effectiveness of PollinateTO in promoting urban biodiversity.
 
 ## Data Features
-The dataset includes several key variables:
+The dataset contains several key variables that are critical for understanding the distribution and characteristics of funded garden projects.
+Below, we describe each variable and its relevance to the analysis:
 
-YEAR_FUNDED: The year the project received funding from PollinateTO.
-GARDEN_TYPE: Categories of gardens (e.g., food gardens, pollinator gardens, rain gardens).
-NIA_OR_EN: Classification of garden location as part of a Neighbourhood Improvement Area (NIA), Emerging Neighborhood (EN), or other areas.
-WARD_NAME: The city ward where the garden is located.
-IS_INDIGENOUS_GARDEN: A binary variable indicating whether the garden is Indigenous-led.
-PROJECT_NAME: Names assigned to specific garden projects.
-LATITUDE and LONGITUDE: Geographic coordinates of each garden.
-These features allow for diverse analyses, from exploring spatial trends to assessing equity and accessibility in funding distribution.
+- YEAR_FUNDED: The year each project received funding from PollinateTO.
+This variable is crucial for understanding temporal trends in the program’s funding allocation.
+It also allows us to analyze shifts in funding priorities over time.
+- GARDEN_TYPE: This variable categorizes the gardens by their ecological functions (e.g., rain gardens, food gardens, pollinator gardens).
+These types reflect the intended ecological role of each garden and are important for understanding the program's impact on biodiversity and the environment.
+- NIA_OR_EN: This classification indicates whether the project is located in a Neighbourhood Improvement Area (NIA), Emerging Neighbourhood (EN), or another area.
+These labels are derived from Toronto's planning data and allow for an exploration of whether funding is disproportionately allocated to areas with different socio-economic characteristics, enabling us to analyze equity in the distribution of funds.
+- WARD_NAME: The specific ward in Toronto where the garden project is located.
+This geographic indicator helps us assess whether certain neighborhoods receive more funding based on local government priorities or demographic characteristics.
+- IS_INDIGENOUS_GARDEN: A binary variable that identifies whether the project is Indigenous-led, based on self-identification in the funding application.
+This is a key variable for exploring whether the program is aligned with Toronto’s commitments to Indigenous communities and ecological knowledge.
+PROJECT_NAME: The name assigned to each garden project.
+While this variable is useful for identification, it is not directly involved in the analysis but helps in linking data points to specific projects.
+- LATITUDE and LONGITUDE: These geographic coordinates provide spatial context for each garden project.
+While these data points help in mapping and visualizing the geographic distribution of funded gardens, their primary role is in understanding spatial patterns and localizing project locations.
 
-## Data Measurement*** (check key)
-The dataset translates real-world community garden projects into structured entries by standardizing key attributes. For example:
+These features collectively allow for a broad analysis of trends, equity, and the ecological goals of the PollinateTO initiative.
 
-YEAR_FUNDED records when projects were initiated, reflecting the temporal dimension of PollinateTO’s funding priorities.
-GARDEN_TYPE is based on project applications, categorized to highlight garden purposes and ecological roles.
-NIA_OR_EN and WARD_NAME capture the geographic and socioeconomic contexts of gardens, derived from city planning frameworks.
-IS_INDIGENOUS_GARDEN identifies projects promoting Indigenous cultural or ecological knowledge, based on project self-identification during funding application.
-The dataset provides structured snapshots of projects, enabling analysis of broader trends, equity impacts, and ecological contributions of urban greening initiatives.
+## Constructed Variables
+In this study, variable is_large_garden was constructed.
+This binary variable was constructed to classify gardens as "large" or "small" based on their estimated garden size.
+Specifically, gardens with an estimated garden size greater than the mean size of all gardens in the dataset were coded as 1 (large garden), while those with a size less than or equal to the mean were coded as 0 (small garden).
+The mean garden size was calculated using the estimated_garden_size variable, excluding any missing values.
+This classification was chosen to capture differences in garden size relative to the overall distribution and to investigate whether larger gardens are more likely to receive funding.
+
+## Data Measurement
+The dataset translates real-world community garden projects funded by PollinateTO into structured entries, enabling the analysis of patterns and relationships.
+It includes several key variables, with the following interpretations based on assumptions about the dataset’s construction:
+
+YEAR_FUNDED likely reflects administrative records of the year projects received funding.
+GARDEN_TYPE seems derived from categorizations in project applications, capturing intended ecological roles.
+NIA_OR_EN appears to use Toronto city planning designations to classify locations.
+IS_INDIGENOUS_GARDEN assumes self-identification during the funding application process.
+WARD_NAME is based on project location within Toronto’s municipal wards.
+
+These assumptions are informed by typical practices in grant-based datasets. Confirmation would require consultation with PollinateTO’s documentation or administrators.
 
@@
 
-Self-Reported Data: Much of the data, including garden types and Indigenous-led status, relies on self-reporting by project applicants.
-Selection Bias: Funding decisions prioritize certain areas (e.g., NIAs), potentially overrepresenting them compared to other regions.
-Temporal Incompleteness: The dataset spans 2019 to present, limiting analyses to recent funding trends.
-Granularity: Certain variables, such as project outcomes or specific plant species, are not included, restricting the scope of ecological impact analysis.
-Geographic Precision: While coordinates are provided, some garden boundaries and precise sizes may not be captured, introducing potential spatial analysis challenges.
+Self-Reported Data: Variables such as GARDEN_TYPE and IS_INDIGENOUS_GARDEN are based on self-reporting by project applicants, which may introduce some subjectivity and inconsistency.
+
+Selection Bias: The dataset reflects funding decisions that may prioritize certain geographic areas, such as NIAs and ENs, which could lead to an overrepresentation of these areas relative to others.
+This potential bias should be considered when analyzing equity and resource allocation.
+
+Temporal Incompleteness: The dataset covers the period from 2019 to the present, which limits our ability to analyze trends over a longer historical period.
+
+Granularity: Some ecological measures, such as specific plant species or detailed environmental outcomes, are not captured in the dataset.
+This restricts the scope of ecological impact analysis, though the dataset’s focus on garden characteristics and funding allocation provides valuable insights.
+
+Geographic Precision: While LATITUDE and LONGITUDE provide the coordinates for each garden, there may be some inaccuracies in the precise location or boundaries of gardens, particularly in cases where they span multiple properties or are irregularly shaped.
 
 ## Methodology
-Several considerations affect the dataset’s completeness and reliability:
-
-Self-Reported Data: Much of the data, including garden types and Indigenous-led status, relies on self-reporting by project applicants.
-Selection Bias: Funding decisions prioritize certain areas (e.g., NIAs), potentially overrepresenting them compared to other regions.
-Temporal Incompleteness: The dataset spans 2019 to present, limiting analyses to recent funding trends.
-Granularity: Certain variables, such as project outcomes or specific plant species, are not included, restricting the scope of ecological impact analysis.
-Geographic Precision: While coordinates are provided, some garden boundaries and precise sizes may not be captured, introducing potential spatial analysis challenges.
+The PollinateTO dataset collects detailed information on community-led garden projects funded by Toronto’s PollinateTO initiative.
+Data is sourced from the city’s Open Data portal and includes variables like garden type, location, funding year, and size, enabling analysis of trends in urban biodiversity projects.
+The dataset uses standardized attributes and is updated annually, making it a valuable tool for evaluating the initiative’s impact across various Toronto neighborhoods, though it may be subject to self-reporting biases and geographical limitations in precision.
 
@@
 In the analysis, I used a Bayesian Logistic Regression Model to examine the relationship between garden size and other factors like the type of garden, the year the project was funded, the neighborhood type (emerging neighborhood, EN vs neighborhood improvement area, NIA), and if it is an indigenous-led project.
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+We are interested in modeling whether a garden is large (binary outcome) as a function of various factors such as the garden type, year funded, neighborhood type, indigenous garden status, and ward name.
+Background details and diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
+```{Latex}
+\begin{align}
+y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{garden_type}_i + \beta_2 \times \mbox{year_funded}_i + \beta_3 \times \mbox{nia_or_en}_i + \beta_4 \times \mbox{is_indigenous_garden}_i + \beta_5 \times \mbox{ward_name}_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta_1 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_2 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_3 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_4 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_5 &\sim \mbox{Normal}(0, 2.5)
+\end{align}
+```
+
 Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
 We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
@@
 Our results are summarized in @tbl-modelresults.
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
 

### 2024-12-03 paper/paper.qmd
Update Paper

@@
 ---
-title: "Key Drivers of Community Garden Growth: Analyzing Trends in PollinateTO Funding Across Toronto’s Neighborhoods"
-subtitle: "A Bayesian Approach to Understanding the Impact of Time, Type, and Location on Garden Size and Number."
+title: "Key Drivers of Community Garden Growth: Analyzing Trends in PollinateTO Funding"
+subtitle: "A Bayesian Approach to Understanding the Impact of Time, Type, and Location on Garden Size and Number in Toronto's Neighborhoods from 2020 to 2023."
 author: 
@@
 number-sections: true
+header-includes:
+  - \usepackage{pdflscape}
 bibliography: references.bib
@@
 library(arrow)
+library(rstanarm)
+library(patchwork)
+library(tinytable)
+library(kableExtra)
+library(broom.mixed)
+
+#### Read data ####
+PT_analysis_data <- read_parquet("../data/02-analysis_data/PT_analysis_data.parquet")
+
+#### Read Model ####
+large_garden_model <- readRDS("../models/large_garden_model.rds")
 ```
@@
 ## Software Packages
-This analysis utilizes the R programming language [@citeR], with the 'rstanarm' [@rstanarm] and 'tidyverse' [@tidyverse] packages for statistical modeling and data manipulation.
+This analysis utilizes the R programming language [@citeR], with the `rstanarm` [@rstanarm] and `tidyverse` [@tidyverse] packages for statistical modeling and data manipulation.
 Our data comes from the PollinateTO Primary Project Garden Locations dataset [@PTdata], which is hosted on Toronto's Open Data Portal.
@@
 The dataset uses standardized attributes and is updated annually, making it a valuable tool for evaluating the initiative’s impact across various Toronto neighborhoods, though it may be subject to self-reporting biases and geographical limitations in precision.
+\newpage
 
@@
 To understand the data provided better, we visualize it using tables and graphs.
-To begin with, let's see how many gardens each type of neighborhood (NIA, EN or None) has and separate them by whether or not they are indigenous-led in @fig-IndNeiBoxPlot and @fig-IndNeiBarChart.
+To begin with, let's see the distribution of Indigenous gardens in each type of neighborhood (NIA, EN or None) in @tbl-IndNei and @fig-IndNeiBarChart.
+As seen in @tbl-IndNei, Indigenous gardens take up a higher proportion of all gardens located in Emerging Neighborhoods; however, there are significantly less Emerging Neighborhood gardens than Neighborhood Improvement Area gardens.
+@fig-IndNeiBarChart helps us realize that overall, there aren't a lot of Indigenous(-led) gardens in this initiative.
+There is only 1 in an Emerging Neighborhood, 1 in a Neighborhood Improvement Area and 3 in other neighborhoods (not EN or NIA).
 
 ```{r}
-#| label: fig-IndNeiBoxPlot
-#| fig-cap: Summary of Gardens by Neighborhood Type
+#| label: tbl-IndNei
+#| tbl-cap: Proportion of Indigenous Gardens in Each Neighborhood Type
 #| echo: false
-
-#### Read data ####
-PT_analysis_data <- read_parquet("../data/02-analysis_data/PT_analysis_data.parquet")
 
@@
     Non_Indigenous_Gardens = Total_Gardens - Indigenous_Gardens, # Count of Non-Indigenous gardens
-    Proportion_Indigenous = (Indigenous_Gardens / Total_Gardens) * 100, # Proportion of Indigenous gardens
+    Proportion_Indigenous = round((Indigenous_Gardens / Total_Gardens) * 100,2), # Proportion of Indigenous gardens
     .groups = "drop"
@@
 #| label: fig-IndNeiBarChart
-#| fig-cap: Distribution of Garden Types Funded Over Time
+#| fig-cap: Number of Indigenous Gardens Funded Across Neighborhoods
 #| echo: false
@@
 ```
-
+\newpage
 Next, we visualize the trend of garden sizes and numbers across time to examine the impact of the initiative over time in @fig-trendline.
+The total number of new gardens funded by PollinateTO decreased significantly from 2020 to 2021, but then increased over 2021 to 2023 by a number greater than the initial drop.
+Estimated Average Garden Size has been decreasing significantly from 2020 to 2022, but recently increase by a small amount from 2022 to 2023.
 
@@
 \newpage
-Now, let's analyze the proportion of each garden type visually across time in @fig-distributiontype to see if certain garden types increased in number and proportion overtime, reflecting shifting funding priorities and initiatives.
+Now, let's analyze the proportion of each garden type across time in @fig-distributiontype to see if certain garden types increased in proportion overtime, reflecting shifting funding priorities and initiatives.
+School Learning Gardens were consistently the most abundant type of garden compared to the rest in each year, and grew in proportion consistently from 2020 to 2023.
+City Park Gardens used to be the second most abundant type of garden in 2020, 2021, and 2022, but they recently fell significantly in proportion and were overtaken by Multiple Front Yard Gardens.
 
@@
 To prevent outliars from making graphs unreadable, we use log-transformed garden sizes to analyze differences in garden sizes based on neighborhood type in @fig-neighborhoodsize.
-A boxplot is used to do this where the 3 neighborhood categories are EN (Emerging Neighborhoods), NIA (Neighborhood Improvement Areas) and None (neither EN nor NIA).
+A dot plot is used to do this where the 3 neighborhood categories are EN (Emerging Neighborhoods), NIA (Neighborhood Improvement Areas) and None (neither EN nor NIA).
 
@@
 #| echo: false
+#| warning: false
+
 # Add a log-transformed column for garden size
@@
 
-# Use the log-transformed column in the boxplot
+# Use the log-transformed column in the dotplot
 ggplot(PT_analysis_data, aes(x = nia_or_en, y = log_garden_size, fill = nia_or_en)) +
-  geom_boxplot(outlier.color = "red", outlier.shape = 16, notch = FALSE) +
+  geom_dotplot(binaxis = "y", stackdir = "center", dotsize = 0.8) +
   scale_fill_brewer(palette = "Pastel1") +
   labs(
+    title = "Log-Transformed Garden Sizes by Neighborhood Type",
     x = "Neighborhood Type (NIA or EN)",
@@
   theme(
-    axis.text.x = element_text(angle = 45, hjust = 1),
-    legend.position = "none"
+    axis.text.x = element_text(angle = 45, hjust = 1)
   )
@@
 
-```{Latex}
 \begin{align} 
-y_i|\pi_i &\sim \mbox{Bern}(\pi_i) \\
-\mbox{logit}(\pi_i) &= \alpha + \beta_1 \times \mbox{garden_type}_i + \beta_2 \times \mbox{year_funded}_i + \beta_3 \times \mbox{nia_or_en}_i + \beta_4 \times \mbox{is_indigenous_garden}_i + \beta_5 \times \mbox{ward_name}_i\\
+y_i|\pi_i & \sim \mbox{Bern}(\pi_i) \\
+\mbox{logit}(\pi_i) &= \alpha 
+    + \beta_1 \times \mbox{garden type}_i 
+    + \beta_2 \times \mbox{year funded}_i \notag \\
+    &\quad + \beta_3 \times \mbox{nia or en}_i 
+    + \beta_4 \times \mbox{is indigenous garden}_i 
+    + \beta_5 \times \mbox{ward name}_i \notag \\
+    &\quad + \beta_6 \times (\mbox{garden type}_i \times \mbox{nia or en}_i) \\
 \alpha &\sim \mbox{Normal}(0, 2.5) \\
@@
 \end{align}
-```
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.  
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+This Bayesian Logistic Regression model predicts whether a garden is classified as "large" ($y_i$), a binary outcome modeled as $y_i \sim \mbox{Bern}(\pi_i)$, where $\pi_i$ is the probability of the garden being large.
+The predictors include: garden type, a categorical variable indicating the type of garden (e.g., pollinator, vegetable); year funded, a continuous variable representing the year the garden received funding; nia or en, a categorical variable indicating whether the garden is in a Neighborhood Improvement Area (NIA) or elsewhere (EN); is indigenous garden, a binary variable identifying whether the garden is indigenous-led; and ward name, a categorical variable representing the geographic ward.
+Additionally, this model includes an interaction term between garden type and nia or en, which helps examine whether the effect of garden type on the likelihood of being large differs depending on whether the garden is in an NIA or EN neighborhood.
+These predictors influence $\pi_i$ through the logit link function: $\mbox{logit}(\pi_i) = \alpha + \beta_1 \times \mbox{garden type}_i + \beta_2 \times \mbox{year funded}_i + \beta_3 \times \mbox{nia or en}_i + \beta_4 \times \mbox{is indigenous garden}_i + \beta_5 \times \mbox{ward name}_i + \beta_6 \times (\mbox{garden type}_i \times \mbox{nia or en}_i)$, where $\alpha$ is the intercept, $\beta_k$ are the coefficients, and $\beta_6$ represents the interaction term between garden type and nia or en.
+Priors for $\alpha$ and $\beta_k$ are defined as Normal(0,2.5), reflecting weakly informative assumptions.
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The Bayesian logistic regression model employed in this analysis is appropriate for investigating the factors influencing the likelihood of a garden being classified as large based on binary outcomes.
+Specifically, this model explores how various predictors—such as garden type, year funded, whether the garden is indigenous-led, the neighborhood classification (NIA or EN), and ward names—affect garden size.
+Given the binary nature of the outcome variable (is_large_garden), logistic regression is the suitable choice as it models probabilities between 0 and 1.
+The inclusion of an interaction term between garden type and nia or en further enriches the analysis, allowing us to examine whether the effect of garden type on garden size differs by neighborhood type.
+This is especially relevant as different neighborhoods may have different conditions and priorities that could influence the type and size of gardens.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The Bayesian framework provides several advantages, including the ability to incorporate prior beliefs about the parameters, produce credible intervals, and assess model uncertainty.
+Furthermore, Bayesian methods enable a more robust assessment of convergence through posterior predictive checks and the examination of diagnostic tools such as trace and Rhat plots, ensuring reliable inferences.
+
+The choice of the Bayesian approach in this context also allows for the flexible inclusion of multiple categorical predictors, which is crucial for this study, as it aims to assess the relationship between categorical factors like garden type, ward names, and the interaction between garden type and nia or en with a quantitative outcome.
+The model's use of informative priors helps regularize the estimates, particularly when dealing with categorical predictors that may have complex relationships with the outcome variable.
+Additionally, the Bayesian methodology provides the opportunity to assess model diagnostics comprehensively, ensuring that the model fits well with the data and that the parameters are estimated with high precision.
+Overall, this model is well-suited to capture the complexities of the data while offering interpretability and flexibility through its probabilistic nature.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-modeltable.
+The Bayesian logistic regression model reveals several important relationships between garden characteristics and the likelihood of a garden being classified as large.
+Among garden types, "Multiple Property Garden" stands out with a substantial positive coefficient (14.94), suggesting that such gardens are strongly associated with larger sizes.
+Conversely, "Boulevard Garden, Rain Garden" exhibits a notable negative coefficient (-13.77), indicating a reduced likelihood of being large.
+Other garden types, such as "Communal Garden" and "School Learning Garden," have smaller coefficients, implying a weaker or negligible relationship with garden size.
+This variability in the influence of garden type highlights the importance of understanding project-specific features when predicting garden outcomes.
+
+The neighborhood context also plays a significant role.
+Gardens located in areas classified as "None" (neither EN nor NIA) have a relatively high positive coefficient (4.54), suggesting that these gardens are more likely to be larger compared to those in ENs or NIAs.
+Additionally, Indigenous-led gardens are associated with a negative coefficient (-8.86), suggesting they may be less likely to be large.
+The interaction terms, such as those between "Multi-Residential Garden" and neighborhood type, indicate that specific combinations of garden type and neighborhood characteristics can further modify the likelihood of a garden being large.
+These findings underscore the complexity of garden size determinants and suggest targeted strategies may be needed to ensure equitable garden funding and development.
 
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-modeltable
 #| warning: false
+#| tbl-cap: "Explanatory model of whether a garden is large based on certain factors: Indigenous-led indicator, Year Funded, Neighborhood, Garden type and Ward Names"
 
-#modelsummary::modelsummary(
-#  list(
-#    "First model" = first_model
- # ),
- # statistic = "mad",
- # fmt = 2
-#)
+# Create the table data
+model_table <- modelsummary::modelsummary(
+  list("Model" = large_garden_model),
+  statistic = "mad",
+  fmt = 2,
+  output = "data.frame" # Output as a data frame for kable compatibility
+)
+
+# Render the table using kableExtra
+model_table %>%
+  kbl() %>%
+  kable_styling(latex_options = c("scale_down", "hold_position")) %>%
+  landscape()
+```
+@fig-model-coefficients visualizes the estimated effects of various predictors on the likelihood of a garden being classified as "large" in the PollinateTO initiative.
+Each point represents the coefficient estimate for a predictor, indicating the direction and strength of its relationship with garden size.
+Predictors with positive coefficients increase the probability of a garden being large, while those with negative coefficients decrease it.
+The horizontal bars show the 95% credible intervals, capturing the uncertainty in the estimates.
+Predictors whose credible intervals do not cross the dashed red line at zero are likely to have significant effects.
+This plot highlights which variables, such as garden type, neighborhood classification, or year funded, are most influential in determining garden size, providing valuable insights into the factors driving outcomes in the program.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-model-coefficients
+#| fig-cap: Coefficients of the Model
+#| fig-height: 8
+
+# Coefficients for the model
+coefficients <- broom.mixed::tidy(large_garden_model, conf.int = TRUE)
+
+# Plot for the coefficients of the model
+ggplot(coefficients, aes(x = estimate, y = term)) +
+  geom_point() + # Coefficient points
+  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high)) + # Confidence/Credible interval bars
+  geom_vline(xintercept = 0, linetype = "dashed", color = "red") + # The zero line
+  labs(x = "Coefficient Estimate", y = "Predictor Variable") +
+  theme_minimal()
+``` 
+\newpage
+## Predictive Probability
+The bar chart in @fig-predictiveprobability visualizes the predicted probability that a garden funded by PollinateTO will be classified as "large," based on the combination of garden type (garden_type) and neighborhood classification (nia_or_en).
+The height of each bar represents the average predicted probability for each garden type and neighborhood combination, with all other variables (such as year funded, Indigenous-led status, and ward) held constant.
+The chart helps identify which garden types and neighborhood categories are most likely to result in larger gardens.
+By examining the differences in predicted probabilities across the garden types and neighborhood types, we can discern patterns in how these factors influence garden size.
+This analysis can inform decisions about resource allocation, funding strategies, or identifying areas where specific types of gardens are more likely to grow larger, contributing to the broader goals of the PollinateTO program.
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+#| label: fig-predictiveprobability
+#| fig-cap: Predictive Probability of Large Gardens
+#| fig-height: 8
+#| fig-width: 16
+#### Predictive Probability Visualization ####
+
+# Load necessary libraries
+library(tidyverse)
+library(ggplot2)
+library(rstanarm)
+
+# Re-factor variables in the dataset
+PT_analysis_data <- PT_analysis_data %>%
+  mutate(
+    garden_type = as.factor(garden_type),
+    nia_or_en = as.factor(nia_or_en),
+    is_indigenous_garden = as.factor(is_indigenous_garden),
+    ward_name = as.factor(ward_name)
+  )
+
+# Generate new data for prediction
+new_data <- expand.grid(
+  garden_type = levels(PT_analysis_data$garden_type),
+  nia_or_en = levels(PT_analysis_data$nia_or_en),
+  year_funded = median(PT_analysis_data$year_funded, na.rm = TRUE),
+  is_indigenous_garden = levels(PT_analysis_data$is_indigenous_garden)[1], # Hold constant
+  ward_name = levels(PT_analysis_data$ward_name)[1] # Hold constant
+)
+
+# Check if new_data is non-empty
+if (nrow(new_data) == 0) stop("new_data is empty. Check factor levels and variables used in expand.grid.")
+
+# Predictive probabilities from the posterior
+predicted_probs <- rstanarm::posterior_predict(large_garden_model, newdata = new_data, type = "response")
+
+# Summarize posterior predictions for each row in new_data
+summary_probs <- as.data.frame(t(predicted_probs)) %>% # Transpose to match new_data rows
+  rowwise() %>%
+  summarise(
+    mean_prob = mean(c_across(everything())),
+    lower_prob = quantile(c_across(everything()), probs = 0.025),
+    upper_prob = quantile(c_across(everything()), probs = 0.975)
+  )
+
+# Combine summary with new_data
+new_data <- bind_cols(new_data, summary_probs)
+
+# Plot the predicted probabilities
+ggplot(new_data, aes(x = garden_type, y = mean_prob, fill = nia_or_en)) +
+  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
+  labs(
+    x = "Garden Type",
+    y = "Predicted Probability",
+    fill = "Neighborhood Type (NIA or EN)"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.y = element_text(size = 14),
+    axis.title.x = element_text(size = 14),
+    axis.text.x = element_text(angle = 45, hjust = 1, size = 12), # Rotates labels by 45 degrees
+    axis.text.y = element_text(size = 12),
+    legend.title = element_text(size = 14),
+    legend.text = element_text(size = 12)   
+  )
 ```
 
+\newpage
 # Discussion
 
-## First discussion point {#sec-first-point}
+## Unearthing Patterns in Garden Allocation {#sec-first-point}
+This paper investigates the factors influencing the size of gardens funded under Toronto’s PollinateTO initiative using a Bayesian logistic regression model.
+Specifically, it explores how garden size (a binary outcome of “large” vs.
+“not large”) is associated with variables such as garden type, year funded, neighborhood classification (Emerging Neighborhoods or Neighborhood Improvement Areas), Indigenous-led status, and ward-level geographic differences.
+By leveraging this model, the analysis identifies patterns and disparities in garden allocation, providing insights into which factors are most predictive of larger gardens.
+The approach incorporates interaction effects to capture nuanced relationships between garden types and neighborhood characteristics.
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The inclusion of credible intervals from the Bayesian model ensures a robust interpretation of uncertainty in coefficient estimates.
+Additionally, key goodness-of-fit measures (e.g., R², WAIC) are evaluated, demonstrating the model’s ability to explain nearly half of the variation in garden size.
+These findings offer a data-driven lens through which urban sustainability programs like PollinateTO can be assessed and improved.
 
-## Second discussion point
+## Neighborhood Inequities in Urban Gardening
 
-Please don't use these as sub-heading labels - change them to be what your point actually is.
+One of the central insights from this analysis is the existence of disparities in garden size across neighborhoods.
+Specifically, gardens in areas classified as “None” (neither EN nor NIA) are significantly more likely to be large, while gardens in NIAs face challenges in achieving similar sizes.
+This suggests that systemic factors, such as socioeconomic inequalities, land availability, and access to resources, influence the outcomes of urban greening initiatives.
 
-## Third discussion point
+This insight highlights a broader inequity in urban sustainability programs: neighborhoods with fewer resources or systemic disadvantages may struggle to achieve comparable environmental benefits.
+This finding emphasizes the importance of embedding equity-focused considerations into grant allocation processes.
+Addressing these disparities can improve not only the environmental impact of such initiatives but also their social relevance and inclusivity.
 
-## Weaknesses and next steps
+## Redefining Success for Indigenous-Led Projects
 
-Weaknesses and next steps should also be included.
+Indigenous-led projects are less likely to result in large gardens, pointing to potential barriers that Indigenous communities face in participating fully in urban greening initiatives.
+This finding underscores the need to reevaluate the criteria for “success” in such programs.
+Indigenous-led projects may prioritize cultural and ecological goals—such as the preservation of native plant species or community education—over garden size.
+
+This raises an important consideration for future urban sustainability programs: qualitative impacts, such as community engagement or cultural significance, may be equally (if not more) important than quantitative measures like garden size.
+By failing to account for these dimensions, existing frameworks risk marginalizing projects that offer critical social and cultural benefits.
+This insight encourages a shift in how success is defined and measured in urban greening efforts.
+
+# Weaknesses and next steps
+I will address some of the weaknesses and limitations of the current analysis, focusing on areas where the methodology or data could be improved.
+Following this, I will explore potential next steps, outlining opportunities for future research and policy development that build on these findings to create more equitable and effective urban sustainability initiatives.
+This section aims to critically reflect on the study while charting a path forward for both academic inquiry and practical applications.
+
+## Challenges and Limitations in Modeling Garden Outcomes
+While the model provides valuable insights, it is not without limitations.
+First, the binary categorization of garden size as “large” vs.
+“not large” may oversimplify the complexity of garden outcomes.
+A more nuanced approach, such as modeling size as a continuous variable or including qualitative measures of success, might provide richer insights.
+
+Second, the dataset is limited to gardens funded by PollinateTO, which may not fully represent the broader urban gardening landscape in Toronto.
+Selection bias could influence the generalizability of these findings to other urban greening initiatives.
+Additionally, the potential for omitted variable bias exists, as factors like community engagement levels or prior gardening experience were not included in the analysis but may significantly impact garden outcomes.
+
+Finally, while interaction terms capture some complexity, they add interpretive challenges, as not all combinations of garden types and neighborhoods had sufficient data to provide precise estimates.
+Future work could address this by gathering larger, more diverse datasets or employing hierarchical models to account for varying group sizes.
+
+## Opportunities for Future Research and Policy Improvements
+Several open questions remain.
+One key area for future research is the inclusion of community-level variables, such as demographic characteristics, access to volunteer networks, and proximity to green spaces.
+These factors may help explain why certain neighborhoods are more likely to host large gardens.
+
+Additionally, further work is needed to explore the qualitative outcomes of urban gardens, particularly for Indigenous-led projects.
+Incorporating interviews, case studies, or community feedback could provide a deeper understanding of how these gardens contribute to cultural preservation and community well-being.
+
+From a methodological perspective, future studies could explore alternative modeling techniques, such as hierarchical Bayesian models, to better account for ward-level variability and cross-neighborhood interactions.
+Expanding the dataset to include other urban greening initiatives or longitudinal data could also provide insights into how garden outcomes evolve over time.
+
+Lastly, there is a need to test the effectiveness of policy interventions informed by these findings.
+For instance, targeted funding for NIAs or revised evaluation criteria for Indigenous-led projects could be piloted and evaluated to determine their impact on garden size and community engagement.
+This iterative process of evidence-based policy design and evaluation can help create more equitable and effective urban sustainability programs.
 
@@
 
+## Data Cleaning
+
+## Data Simulation
+
 # Model details {#sec-model-details}
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+In @fig-ppcheck we implement a posterior predictive check.
+The posterior predictive check (pp_check) helps assess how well the model fits the observed data in the context of your research on the factors influencing community garden funding by PollinateTO across different Toronto neighborhoods.
+By comparing the model’s predictions (posterior distribution) with the observed data, this diagnostic reveals whether the model is accurately capturing the key influences, such as garden type, funding year, and neighborhood characteristics.
+A good fit suggests that these factors are well-represented in the model, while significant deviations between the predictions and the data may indicate that the model needs refinement.
+This check is essential for validating the model’s effectiveness in answering your research question.
 
 ```{r}
+#| label: fig-ppcheck
 #| eval: true
@@
 #| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
 #| fig-cap: "Examining how the model fits, and is affected by, the data"
@@
 
-#pp_check(first_model) +
-#  theme_classic() +
-#  theme(legend.position = "bottom")
+pp_check(large_garden_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+```
+\newpage
+In @fig-posteriorvsprior we compare the posterior with the prior.
+The posterior vs.
+prior comparison visualizes how the model’s posterior estimates for various parameters compare to the prior distributions.
+In the context of your research on the factors influencing the size and number of community gardens funded by PollinateTO, this comparison helps evaluate whether the data has significantly updated the prior beliefs embedded in the model.
+A well-calibrated model should show a substantial difference between the prior and posterior for key parameters, indicating that the observed data has influenced the model’s estimates.
+This comparison is crucial for understanding which variables are most influential in determining garden size and funding across neighborhoods.
 
-#posterior_vs_prior(first_model) +
-#  theme_minimal() +
- # scale_color_brewer(palette = "Set1") +
-#  theme(legend.position = "bottom") +
-#  coord_flip()
+Given the large number of parameters in the model, a random sample of 5 garden types and 5 ward names was taken for this visualization.
+This reduces complexity and ensures the comparison remains interpretable, focusing on a manageable subset of the factors that may affect the size and number of community gardens across Toronto.
+
+```{r}
+#| label: fig-posteriorvsprior
+#| fig-cap: Compare the posterior with the prior 
+#| eval: true
+#| echo: false
+#| warning: false
+#| message: false
+
+PT_analysis_data <- PT_analysis_data %>%
+mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0))
+
+garden_type_list <- c("garden_typeBoulevard Garden, Rain Garden","garden_typeCity Park","garden_typeCity Park, Rain Garden","garden_typeCommunal Garden","garden_typeMulti-Residential Garden","garden_typeMultiple Boulevard Gardens","garden_typeMultiple Front Yard Gardens","garden_typeMultiple Property Garden","garden_typeSchool Learning Garden","garden_typeSpiritual Centre Garden","garden_typeSpiritual Centre Garden, Rain Garden")
+set.seed(522)
+garden_type_sample <- sample(garden_type_list,5)
+
+ward_names_list <- c("ward_nameDavenport","ward_nameDon Valley East","ward_nameDon Valley North","ward_nameDon Valley West","ward_nameEglinton-Lawrence","ward_nameEtobicoke Centre","ward_nameEtobicoke North","ward_nameEtobicoke-Lakeshore","ward_nameEtobicoke-North","ward_nameHumber River-Black Creek","ward_nameParkdale-High Park","ward_nameScarborough Centre","ward_nameScarborough Southwest","ward_nameScarborough-Agincourt","ward_nameScarborough-Guildwood","ward_nameScarborough-Rouge Park","ward_nameSpadina-Fort York","ward_nameToronto Centre","ward_nameToronto-Danforth","ward_nameToronto-St.
+Paul's","ward_nameUniversity-Rosedale","ward_nameWillowdale","ward_nameYork Centre","ward_nameYork South-Weston")
+set.seed(522)
+ward_names_sample <- sample(ward_names_list,5)
+
+pars_to_plot <- c("(Intercept)","is_indigenous_garden1","year_funded","nia_or_enNIA",  "nia_or_enNone", garden_type_sample, ward_names_sample)
+
+# Compare posterior and prior
+posterior_vs_prior(large_garden_model,
+                   pars = pars_to_plot, size = 0.2) +
+  theme_minimal() +
+  theme(legend.position = "bottom", # Position the legend at the bottom
+        legend.key.size = unit(0.05, 'cm'),
+        legend.spacing.x = unit(0.1, 'cm'), # Reduce horizontal spacing between legend items
+        axis.text.x = element_text(hjust = 1, vjust = 0.5), # Adjust x-axis text
+        text = element_text(size = 8) # Reduce overall text size
+  ) +
+  coord_flip()
 ```
 
-## Diagnostics
+## Markov Chain Monte Carlo Convergence Check
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+@fig-traceplots displays trace plots of the model.
+Trace plots are essential diagnostics in Bayesian logistic regression models, like the one used in your research examining predictors of whether a garden is larger than average, including garden type, indigenous-led status, funding year, and location.
+These plots display the sampled parameter values across MCMC iterations, helping assess convergence and mixing.
+A well-behaved trace plot resembles a "fuzzy caterpillar," indicating that the chain has stabilized and is sampling consistently from the posterior distribution, ensuring reliable estimates.
+This matches the description of what we observe in our trace plots.
+Non-convergence, shown by trends or failure to stabilize, may result from poor model specification or insufficient warmup.
+Good mixing is characterized by rapid oscillations and variability, reflecting effective exploration of the parameter space, whereas slow mixing or long stretches of similar values may indicate inefficiency or high autocorrelation.
+Proper convergence and mixing are critical to draw robust inferences about how factors like indigenous-led status or funding year influence garden size, ensuring the validity of your findings.
+This suggests...
 
@@
 #| warning: false
-#| label: fig-stanareyouokay
+#| label: fig-traceplots
 #| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
+#| fig-subcap: Trace plot of Intercept, Indigenous-led indicator, Year Funded, Neighborhood, Garden type (Sample) and Ward Names (Sample)
+#| fig-width: 10
+#| fig-height: 11
 
-#plot(first_model, "trace")
+# Generate individual plots
+plot1 <- plot(large_garden_model, "trace", "(Intercept)")
+plot2 <- plot(large_garden_model, "trace", "is_indigenous_garden1")
+plot3 <- plot(large_garden_model, "trace", "year_funded")
+plot4 <- plot(large_garden_model, "trace", "nia_or_enNIA")
+plot5 <- plot(large_garden_model, "trace", "nia_or_enNone")
+plot6 <- plot(large_garden_model, "trace", garden_type_sample[1])
+plot7 <- plot(large_garden_model, "trace", garden_type_sample[2])
+plot8 <- plot(large_garden_model, "trace", garden_type_sample[3])
+plot9 <- plot(large_garden_model, "trace", garden_type_sample[4])
+plot10 <- plot(large_garden_model, "trace", garden_type_sample[5])
+plot11 <- plot(large_garden_model, "trace", ward_names_sample[1])
+plot12 <- plot(large_garden_model, "trace", ward_names_sample[2])
+plot13 <- plot(large_garden_model, "trace", ward_names_sample[3])
+plot14 <- plot(large_garden_model, "trace", ward_names_sample[4])
+plot15 <- plot(large_garden_model, "trace", ward_names_sample[5])
 
-#plot(first_model, "rhat")
+# Combine the plots side by side (2 rows)
+combined_plot <- (plot1 | plot2 | plot3) / (plot4 | plot5 | plot6) / (plot7 | plot8 | plot9) / (plot10 | plot11 | plot12) / (plot13 | plot14 | plot15)
+
+# Display combined plot
+combined_plot
+```
+
+@fig-rhatplot is a Rhat plot.
+Rhat plots are critical diagnostics for assessing convergence in Bayesian logistic regression models like this one.
+The Rhat statistic, or Gelman-Rubin diagnostic, compares within-chain and between-chain variance to determine if the Markov chains have converged to the posterior distribution.
+Ideally, Rhat values should be close to 1 (e.g., below 1.01 or 1.05), indicating convergence.
+This matches the description of what we observe in our Rhat plot.
+An Rhat plot visually summarizes these values, with deviations above the threshold signaling potential non-convergence and the need for more iterations, better model specification, or improved priors.
+By ensuring all parameters in the model have Rhat values near 1, the plot provides confidence in the stability and reliability of the posterior estimates.
+This is crucial for interpreting how predictors like indigenous-led status and funding year influence garden size accurately.
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-rhatplot
+#| fig-cap: "Rhat plot"
+#| fig-height: 10
+
+plot(large_garden_model, "rhat")
 ```

### 2024-12-03 scripts/03-clean_data.R
Delete scripts directory

@@ -1,54 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-## DELETE COLUMNS IM NOT USING AND RELEVANT CLEANING
-
-#### Workspace setup ####
-library(tidyverse)
-library(dplyr)
-library(arrow)
-
-#### Clean data ####
-PT_raw_data <- read_csv("./data/01-raw_data/PT_raw_data.csv")
-
-cleaned_data <-
-  PT_raw_data |>
-  janitor::clean_names() |>
-  # Remove the redundant 'x_id' column as it is same as 'id' column
-  select(-x_id) |>
-  # Remove columns that have no data i.e. all values are "None"
-  select (-description) |>
-  select (-image_name) |>
-  select (-image_alt_text) |>
-  # ensure the year funded, number of gardens, estimated garden size are numbers
-  mutate(
-    year_funded = as.numeric(year_funded),
-    number_of_gardens = as.numeric(number_of_gardens),
-    estimated_garden_size = as.numeric(estimated_garden_size),
-    # ensure the neighborhood number, ward_number, id, nia_or_en, group name, project name, garden type, location name, garden address, postal code, ward name and geometry are strings
-    neighbourhood_number = as.character(neighbourhood_number),
-    ward_number = as.character(ward_number),
-    id = as.character(id),
-    nia_or_en = as.character(nia_or_en),
-    group_name = as.character(group_name),
-    project_name = as.character(project_name),
-    garden_type = as.character(garden_type),
-    primary_location_name = as.character(primary_location_name),
-    primary_garden_address = as.character(primary_garden_address),
-    postal_code = as.character(postal_code),
-    ward_name = as.character(ward_name),
-    geometry = as.character(geometry),
-    # Convert the is_indigenous_garden column to binary column where 1 means that the project is indigenous led
-    is_indigenous_garden = if_else(is_indigenous_garden == "Y", 1, 0)
-  ) |>
-  #drop any rows with missing/NA values
-  tidyr::drop_na()
-
[…]

### 2024-12-03 scripts/04-test_analysis_data.R

@@ -1,69 +1,0 @@
-#### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(testthat)
-
-data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
-})
-
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
-})
-
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
-})
-
-# Test that the 'party' column is character type
-test_that("'party' is character", {
-  expect_type(analysis_data$party, "character")
-})
-
-# Test that the 'state' column is character type
-test_that("'state' is character", {
-  expect_type(analysis_data$state, "character")
-})
-
-# Test that there are no missing values in the dataset
-test_that("no missing values in dataset", {
-  expect_true(all(!is.na(analysis_data)))
-})
-
-# Test that 'division' contains unique values (no duplicates)
-test_that("'division' column contains unique values", {
-  expect_equal(length(unique(analysis_data$division)), 151)
-})
-
-# Test that 'state' contains only valid Australian state or territory names
-valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
-                  "Tasmania", "Northern Territory", "Australian Capital Territory")
-test_that("'state' contains valid Australian state names", {
-  expect_true(all(analysis_data$state %in% valid_states))
-})
-
-# Test that there are no empty strings in 'division', 'party', or 'state' columns
-test_that("no empty strings in 'division', 'party', or 'state' columns", {
-  expect_false(any(analysis_data$division == "" | analysis_data$party == "" | analysis_data$state == ""))
-})
-
-# Test that the 'party' column contains at least 2 unique values
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,150 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(ggplot2)
-library(dplyr)
-library(knitr)
-
-#### Read data ####
-PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-# Summarize the data by year to get the total number of gardens and average garden size per year
-PT_summary <- PT_analysis_data %>%
-  group_by(year_funded) %>%
-  summarise(
-    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
-    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
-  )
-
-# Create the line chart with two y-axes (total gardens and average size)
-ggplot(PT_summary, aes(x = year_funded)) +
-  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
-  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
-  scale_y_continuous(
-    name = "Total Number of Gardens", 
-    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
-  ) +
-  labs(
-    title = "Trend of Garden Size and Number Over Time",
-    x = "Year Funded",
-    y = "Total Number of Gardens"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.title.y.right = element_text(color = "red"),
-    axis.text.y.right = element_text(color = "red")
-  )
-
-# Summarize the data to get the count of garden types by year
-PT_summary <- PT_analysis_data %>%
-  group_by(year_funded, garden_type) %>%
-  summarise(
-    garden_count = n(), # Count the number of occurrences for each garden type per year
-    .groups = "drop"
-  )
-
-# Create the bar chart
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,71 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(brms)
-library(arrow)
-
-# Load dataset
-PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-# Prepare the data
-PT_analysis_data <- PT_analysis_data %>%
-  # Create a binary outcome variable for whether the garden is larger than average
-  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>%
-  # Ensure categorical variables are treated as factors
-  mutate(
-    garden_type = as.factor(garden_type),
-    ward_name = as.factor(ward_name),
-    is_indigenous_garden = as.factor(is_indigenous_garden),
-    nia_or_en = as.factor(nia_or_en),
-    year_funded = as.numeric(year_funded) # Treated as continuous
-  )
-
-# Fit a Bayesian logistic regression model
-bayesian_model <- brm(
-  formula = is_large_garden ~ garden_type + year_funded + nia_or_en + is_indigenous_garden + ward_name,
-  family = bernoulli(link = "logit"), # Logistic regression for binary outcomes
-  data = PT_analysis_data,
-  prior = c(
-    prior(normal(0, 1), class = "b"),   # Priors for coefficients
-    prior(cauchy(0, 2), class = "Intercept") # Prior for intercept
-  ),
-  iter = 2000,  # Number of iterations for MCMC
-  warmup = 500, # Warmup period for MCMC
-  chains = 4,   # Number of MCMC chains
-  cores = 4,    # Number of cores for parallel computation
-  seed = 123    # Random seed for reproducibility
-)
-
-# Print a summary of the model
-summary(bayesian_model)
-
-# Plot posterior distributions of coefficients
-plot(bayesian_model)
-
-# Posterior summaries for all predictors
-posterior_summary(bayesian_model)
-
[…]

### 2024-12-03 scripts/02-download_data.R
Update scripts

@@ -1,0 +1,23 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+#### Download data ####
+package <- show_package("pollinateto-primary-project-garden-locations")
+resources <- list_package_resources("pollinateto-primary-project-garden-locations")
+csv_resource <- resources[resources$format == "CSV", ]
+first_csv_resource <- csv_resource[1, ]
+csv_id <- first_csv_resource$id
+pollinateto_data <- get_resource(csv_id) 
+
+#### Save data ####
+write_csv(pollinateto_data, "./data/01-raw_data/PT_raw_data.csv")

### 2024-12-03 scripts/03-clean_data.R

@@ -1,0 +1,54 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+## DELETE COLUMNS IM NOT USING AND RELEVANT CLEANING
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(arrow)
+
+#### Clean data ####
+PT_raw_data <- read_csv("./data/01-raw_data/PT_raw_data.csv")
+
+cleaned_data <-
+  PT_raw_data |>
+  janitor::clean_names() |>
+  # Remove the redundant 'x_id' column as it is same as 'id' column
+  select(-x_id) |>
+  # Remove columns that have no data i.e. all values are "None"
+  select (-description) |>
+  select (-image_name) |>
+  select (-image_alt_text) |>
+  # ensure the year funded, number of gardens, estimated garden size are numbers
+  mutate(
+    year_funded = as.numeric(year_funded),
+    number_of_gardens = as.numeric(number_of_gardens),
+    estimated_garden_size = as.numeric(estimated_garden_size),
+    # ensure the neighborhood number, ward_number, id, nia_or_en, group name, project name, garden type, location name, garden address, postal code, ward name and geometry are strings
+    neighbourhood_number = as.character(neighbourhood_number),
+    ward_number = as.character(ward_number),
+    id = as.character(id),
+    nia_or_en = as.character(nia_or_en),
+    group_name = as.character(group_name),
+    project_name = as.character(project_name),
+    garden_type = as.character(garden_type),
+    primary_location_name = as.character(primary_location_name),
+    primary_garden_address = as.character(primary_garden_address),
+    postal_code = as.character(postal_code),
+    ward_name = as.character(ward_name),
+    geometry = as.character(geometry),
+    # Convert the is_indigenous_garden column to binary column where 1 means that the project is indigenous led
+    is_indigenous_garden = if_else(is_indigenous_garden == "Y", 1, 0)
+  ) |>
+  #drop any rows with missing/NA values
+  tidyr::drop_na()
+
[…]

### 2024-12-03 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,149 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(arrow)
+
+#### Read data ####
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Summarize the data by year to get the total number of gardens and average garden size per year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded) %>%
+  summarise(
+    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
+    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
+  )
+
+# Create the line chart with two y-axes (total gardens and average size)
+ggplot(PT_summary, aes(x = year_funded)) +
+  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
+  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
+  scale_y_continuous(
+    name = "Total Number of Gardens", 
+    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
+  ) +
+  labs(
+    title = "Trend of Garden Size and Number Over Time",
+    x = "Year Funded",
+    y = "Total Number of Gardens"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "red"),
+    axis.text.y.right = element_text(color = "red")
+  )
+
+# Summarize the data to get the count of garden types by year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded, garden_type) %>%
+  summarise(
+    garden_count = n(), # Count the number of occurrences for each garden type per year
+    .groups = "drop"
+  )
+
+# Create the bar chart
[…]

### 2024-12-03 scripts/06-model_data.R

@@ -1,0 +1,47 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(arrow)
+
+# Load dataset
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+
+# Prepare the data
+PT_analysis_data <- PT_analysis_data %>%
+  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>% # In binary form
+  mutate(
+    garden_type = as.factor(garden_type),
+    ward_name = as.factor(ward_name),
+    is_indigenous_garden = as.factor(is_indigenous_garden),
+    nia_or_en = as.factor(nia_or_en),
+    year_funded = as.numeric(year_funded) # Treated as continuous
+  )
+
+# Fit a Bayesian logistic regression model using rstanarm
+large_garden_model <- stan_glm(
+  is_large_garden ~ garden_type * nia_or_en + year_funded + is_indigenous_garden + ward_name, # Adding interaction between garden_type and nia_or_en
+  family = binomial(link = "logit"),
+  data = PT_analysis_data,
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger normal priors
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger prior on intercept
+  prior_aux = exponential(rate = 1),  # Regularizing the auxiliary parameters (error terms)
+  seed = 522,
+  chains = 4,  # Ensure more sampling chains
+  iter = 2000  # Set enough iterations for convergence
+)
+
+#### Save model ####
+saveRDS(
+  large_garden_model,
+  file = "./models/large_garden_model.rds"
+)

### 2024-12-03 scripts/06-model_data_working.R

@@ -1,0 +1,70 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+#### Workspace setup ####
+library(tidyverse)
+library(brms)
+library(arrow)
+
+# Load dataset
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Prepare the data
+PT_analysis_data <- PT_analysis_data %>%
+  # Create a binary outcome variable for whether the garden is larger than average
+  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>%
+  # Ensure categorical variables are treated as factors
+  mutate(
+    garden_type = as.factor(garden_type),
+    ward_name = as.factor(ward_name),
+    is_indigenous_garden = as.factor(is_indigenous_garden),
+    nia_or_en = as.factor(nia_or_en),
+    year_funded = as.numeric(year_funded) # Treated as continuous
+  )
+
+# Fit a Bayesian logistic regression model
+bayesian_model <- brm(
+  formula = is_large_garden ~ garden_type + year_funded + nia_or_en + is_indigenous_garden + ward_name,
+  family = bernoulli(link = "logit"), # Logistic regression for binary outcomes
+  data = PT_analysis_data,
+  prior = c(
+    prior(normal(0, 1), class = "b"),   # Priors for coefficients
+    prior(cauchy(0, 2), class = "Intercept") # Prior for intercept
+  ),
+  iter = 2000,  # Number of iterations for MCMC
+  warmup = 500, # Warmup period for MCMC
+  chains = 4,   # Number of MCMC chains
+  cores = 4,    # Number of cores for parallel computation
+  seed = 123    # Random seed for reproducibility
+)
+
+# Print a summary of the model
+summary(bayesian_model)
+
+# Plot posterior distributions of coefficients
+plot(bayesian_model)
+
+# Posterior summaries for all predictors
+posterior_summary(bayesian_model)
+
[…]

### 2024-12-03 scripts/06-model_data_working.R
Delete scripts/06-model_data_working.R

@@ -1,70 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Workspace setup ####
-library(tidyverse)
-library(brms)
-library(arrow)
-
-# Load dataset
-PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-# Prepare the data
-PT_analysis_data <- PT_analysis_data %>%
-  # Create a binary outcome variable for whether the garden is larger than average
-  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>%
-  # Ensure categorical variables are treated as factors
-  mutate(
-    garden_type = as.factor(garden_type),
-    ward_name = as.factor(ward_name),
-    is_indigenous_garden = as.factor(is_indigenous_garden),
-    nia_or_en = as.factor(nia_or_en),
-    year_funded = as.numeric(year_funded) # Treated as continuous
-  )
-
-# Fit a Bayesian logistic regression model
-bayesian_model <- brm(
-  formula = is_large_garden ~ garden_type + year_funded + nia_or_en + is_indigenous_garden + ward_name,
-  family = bernoulli(link = "logit"), # Logistic regression for binary outcomes
-  data = PT_analysis_data,
-  prior = c(
-    prior(normal(0, 1), class = "b"),   # Priors for coefficients
-    prior(cauchy(0, 2), class = "Intercept") # Prior for intercept
-  ),
-  iter = 2000,  # Number of iterations for MCMC
-  warmup = 500, # Warmup period for MCMC
-  chains = 4,   # Number of MCMC chains
-  cores = 4,    # Number of cores for parallel computation
-  seed = 123    # Random seed for reproducibility
-)
-
-# Print a summary of the model
-summary(bayesian_model)
-
-# Plot posterior distributions of coefficients
-plot(bayesian_model)
-
-# Posterior summaries for all predictors
-posterior_summary(bayesian_model)
-
[…]

### 2024-12-03 scripts/00-simulate_data.R
Update Simulate and Clean scripts

@@ -1,0 +1,86 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+  #state and party that won each division.
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `starter_folder` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(522)
+
+
+#### Simulate data ####
+
+#### Define years of funding ####
+years_of_funding <- c(2020,2021,2022,2023)
+
+#### Garden types list ####
+garden_types <- c(
+  "Multiple Front Yard Gardens",
+  "Communal Garden",
+  "School Learning Garden",
+  "City Park",
+  "Spiritual Centre Garden",
+  "Boulevard Garden, Rain Garden",
+  "Multi-Residential Garden",
+  "Spiritual Centre Garden, Rain Garden",
+  "Boulevard Garden",
+  "City Park, Rain Garden",
+  "Multiple Boulevard Gardens",
+  "Multiple Property Garden"
+)
+
+nia_or_en_options <- c("NIA","EN","None")
+
+ward_names <- c("Davenport","Scarborough-Agincourt","Toronto-Danforth",
+                "Etobicoke Centre","Scarborough Southwest",
+                "Etobicoke-Lakeshore","University-Rosedale","Toronto-St. Paul's",
+                "Toronto Centre","Don Valley East","Eglinton-Lawrence","Parkdale-High Park",
+                "Beaches-East York","Scarborough-Rouge Park","Humber River-Black Creek",
+                "Spadina-Fort York","York Centre","Etobicoke North","Scarborough-Guildwood",
+                "York South-Weston","Willowdale","Don Valley West",
+                "Don Valley North","Scarborough Centre")
+
+is_indigenous_garden_options <- c(0,1)
+
+# Create a dataset by randomly assigning values to variables for each id
+simulated_data <- tibble(
+  id = paste(1:149),
+  year_funded = sample(
+    years_of_funding,
+    size = 149,
+    replace = TRUE
+  ),
+  garden_type = sample(
+    garden_types,
+    size = 149,
+    replace = TRUE
+  ),
+  nia_or_en = sample(
+    nia_or_en_options,
+    size = 149,
[…]

### 2024-12-03 scripts/03-clean_data.R

@@ -1,0 +1,57 @@
+#### Preamble ####
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 6 April 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+## DELETE COLUMNS IM NOT USING AND RELEVANT CLEANING
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(arrow)
+
+#### Clean data ####
+PT_raw_data <- read_csv("./data/01-raw_data/PT_raw_data.csv")
+
+cleaned_data <-
+  PT_raw_data |>
+  janitor::clean_names() |>
+  # Remove the redundant 'x_id' column as it is same as 'id' column
+  select(-x_id) |>
+  # Remove columns that have no data i.e. all values are "None"
+  select (-description) |>
+  select (-image_name) |>
+  select (-image_alt_text) |>
+  # Remove unused columns
+  select (-group_name) |>
+  select (-project_name) |>
+  select (-neighbourhood_number) |>
+  select (-primary_location_name) |>
+  select (-primary_garden_address) |>
+  select (-postal_code) |>
+  select (-geometry) |>
+  select (-ward_number) |>
+  # ensure the year funded, number of gardens, estimated garden size are numbers
+  mutate(
+    year_funded = as.numeric(year_funded),
+    number_of_gardens = as.numeric(number_of_gardens),
+    estimated_garden_size = as.numeric(estimated_garden_size),
+    # ensure the id, nia_or_en, garden type, and ward name are strings
+    id = as.character(id),
+    nia_or_en = as.character(nia_or_en),
+    garden_type = as.character(garden_type),
+    ward_name = as.character(ward_name),
+    # Convert the is_indigenous_garden column to binary column where 1 means that the project is indigenous led
+    is_indigenous_garden = if_else(is_indigenous_garden == "Y", 1, 0),
+    ward_name = if_else(ward_name == "Etobicoke-North", "Etobicoke North", ward_name)
+  ) |>
+  #drop any rows with missing/NA values
+  tidyr::drop_na()
+
+#### Save data ####
+write_parquet(cleaned_data, "./data/02-analysis_data/PT_analysis_data.parquet")
+

### 2024-12-04 scripts/00-simulate_data.R
Delete scripts directory

@@ -1,86 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
-
-
-#### Workspace setup ####
-library(tidyverse)
-set.seed(522)
-
-
-#### Simulate data ####
-
-#### Define years of funding ####
-years_of_funding <- c(2020,2021,2022,2023)
-
-#### Garden types list ####
-garden_types <- c(
-  "Multiple Front Yard Gardens",
-  "Communal Garden",
-  "School Learning Garden",
-  "City Park",
-  "Spiritual Centre Garden",
-  "Boulevard Garden, Rain Garden",
-  "Multi-Residential Garden",
-  "Spiritual Centre Garden, Rain Garden",
-  "Boulevard Garden",
-  "City Park, Rain Garden",
-  "Multiple Boulevard Gardens",
-  "Multiple Property Garden"
-)
-
-nia_or_en_options <- c("NIA","EN","None")
-
-ward_names <- c("Davenport","Scarborough-Agincourt","Toronto-Danforth",
-                "Etobicoke Centre","Scarborough Southwest",
-                "Etobicoke-Lakeshore","University-Rosedale","Toronto-St. Paul's",
-                "Toronto Centre","Don Valley East","Eglinton-Lawrence","Parkdale-High Park",
-                "Beaches-East York","Scarborough-Rouge Park","Humber River-Black Creek",
-                "Spadina-Fort York","York Centre","Etobicoke North","Scarborough-Guildwood",
-                "York South-Weston","Willowdale","Don Valley West",
-                "Don Valley North","Scarborough Centre")
-
-is_indigenous_garden_options <- c(0,1)
-
-# Create a dataset by randomly assigning values to variables for each id
-simulated_data <- tibble(
-  id = paste(1:149),
-  year_funded = sample(
-    years_of_funding,
-    size = 149,
-    replace = TRUE
-  ),
-  garden_type = sample(
-    garden_types,
-    size = 149,
-    replace = TRUE
-  ),
-  nia_or_en = sample(
-    nia_or_en_options,
-    size = 149,
[…]

### 2024-12-04 scripts/02-download_data.R

@@ -1,23 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-
-#### Download data ####
-package <- show_package("pollinateto-primary-project-garden-locations")
-resources <- list_package_resources("pollinateto-primary-project-garden-locations")
-csv_resource <- resources[resources$format == "CSV", ]
-first_csv_resource <- csv_resource[1, ]
-csv_id <- first_csv_resource$id
-pollinateto_data <- get_resource(csv_id) 
-
-#### Save data ####
-write_csv(pollinateto_data, "./data/01-raw_data/PT_raw_data.csv")

### 2024-12-04 scripts/03-clean_data.R

@@ -1,57 +1,0 @@
-#### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-## DELETE COLUMNS IM NOT USING AND RELEVANT CLEANING
-
-#### Workspace setup ####
-library(tidyverse)
-library(dplyr)
-library(arrow)
-
-#### Clean data ####
-PT_raw_data <- read_csv("./data/01-raw_data/PT_raw_data.csv")
-
-cleaned_data <-
-  PT_raw_data |>
-  janitor::clean_names() |>
-  # Remove the redundant 'x_id' column as it is same as 'id' column
-  select(-x_id) |>
-  # Remove columns that have no data i.e. all values are "None"
-  select (-description) |>
-  select (-image_name) |>
-  select (-image_alt_text) |>
-  # Remove unused columns
-  select (-group_name) |>
-  select (-project_name) |>
-  select (-neighbourhood_number) |>
-  select (-primary_location_name) |>
-  select (-primary_garden_address) |>
-  select (-postal_code) |>
-  select (-geometry) |>
-  select (-ward_number) |>
-  # ensure the year funded, number of gardens, estimated garden size are numbers
-  mutate(
-    year_funded = as.numeric(year_funded),
-    number_of_gardens = as.numeric(number_of_gardens),
-    estimated_garden_size = as.numeric(estimated_garden_size),
-    # ensure the id, nia_or_en, garden type, and ward name are strings
-    id = as.character(id),
-    nia_or_en = as.character(nia_or_en),
-    garden_type = as.character(garden_type),
-    ward_name = as.character(ward_name),
-    # Convert the is_indigenous_garden column to binary column where 1 means that the project is indigenous led
-    is_indigenous_garden = if_else(is_indigenous_garden == "Y", 1, 0),
-    ward_name = if_else(ward_name == "Etobicoke-North", "Etobicoke North", ward_name)
-  ) |>
-  #drop any rows with missing/NA values
-  tidyr::drop_na()
-
-#### Save data ####
-write_parquet(cleaned_data, "./data/02-analysis_data/PT_analysis_data.parquet")
-

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -1,149 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(ggplot2)
-library(dplyr)
-library(knitr)
-library(arrow)
-
-#### Read data ####
-PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-# Summarize the data by year to get the total number of gardens and average garden size per year
-PT_summary <- PT_analysis_data %>%
-  group_by(year_funded) %>%
-  summarise(
-    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
-    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
-  )
-
-# Create the line chart with two y-axes (total gardens and average size)
-ggplot(PT_summary, aes(x = year_funded)) +
-  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
-  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
-  scale_y_continuous(
-    name = "Total Number of Gardens", 
-    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
-  ) +
-  labs(
-    title = "Trend of Garden Size and Number Over Time",
-    x = "Year Funded",
-    y = "Total Number of Gardens"
-  ) +
-  theme_minimal() +
-  theme(
-    axis.title.y.right = element_text(color = "red"),
-    axis.text.y.right = element_text(color = "red")
-  )
-
-# Summarize the data to get the count of garden types by year
-PT_summary <- PT_analysis_data %>%
-  group_by(year_funded, garden_type) %>%
-  summarise(
-    garden_count = n(), # Count the number of occurrences for each garden type per year
-    .groups = "drop"
-  )
-
-# Create the bar chart
[…]

### 2024-12-04 scripts/06-model_data.R

@@ -1,47 +1,0 @@
-#### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-#### Workspace setup ####
-library(tidyverse)
-library(rstanarm)
-library(arrow)
-
-# Load dataset
-PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
-
-
-# Prepare the data
-PT_analysis_data <- PT_analysis_data %>%
-  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>% # In binary form
-  mutate(
-    garden_type = as.factor(garden_type),
-    ward_name = as.factor(ward_name),
-    is_indigenous_garden = as.factor(is_indigenous_garden),
-    nia_or_en = as.factor(nia_or_en),
-    year_funded = as.numeric(year_funded) # Treated as continuous
-  )
-
-# Fit a Bayesian logistic regression model using rstanarm
-large_garden_model <- stan_glm(
-  is_large_garden ~ garden_type * nia_or_en + year_funded + is_indigenous_garden + ward_name, # Adding interaction between garden_type and nia_or_en
-  family = binomial(link = "logit"),
-  data = PT_analysis_data,
-  prior = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger normal priors
-  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger prior on intercept
-  prior_aux = exponential(rate = 1),  # Regularizing the auxiliary parameters (error terms)
-  seed = 522,
-  chains = 4,  # Ensure more sampling chains
-  iter = 2000  # Set enough iterations for convergence
-)
-
-#### Save model ####
-saveRDS(
-  large_garden_model,
-  file = "./models/large_garden_model.rds"
-)

### 2024-12-04 scripts/00-simulate_data.R
Add files via upload

@@ -1,0 +1,86 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of PollinateTO Primary Project Garden Locations.
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+set.seed(522)
+
+
+#### Simulate data ####
+
+#### Define years of funding ####
+years_of_funding <- c(2020,2021,2022,2023)
+
+#### Garden types list ####
+garden_types <- c(
+  "Multiple Front Yard Gardens",
+  "Communal Garden",
+  "School Learning Garden",
+  "City Park",
+  "Spiritual Centre Garden",
+  "Boulevard Garden, Rain Garden",
+  "Multi-Residential Garden",
+  "Spiritual Centre Garden, Rain Garden",
+  "Boulevard Garden",
+  "City Park, Rain Garden",
+  "Multiple Boulevard Gardens",
+  "Multiple Property Garden"
+)
+
+nia_or_en_options <- c("NIA","EN","None")
+
+ward_names <- c("Davenport","Scarborough-Agincourt","Toronto-Danforth",
+                "Etobicoke Centre","Scarborough Southwest",
+                "Etobicoke-Lakeshore","University-Rosedale","Toronto-St. Paul's",
+                "Toronto Centre","Don Valley East","Eglinton-Lawrence","Parkdale-High Park",
+                "Beaches-East York","Scarborough-Rouge Park","Humber River-Black Creek",
+                "Spadina-Fort York","York Centre","Etobicoke North","Scarborough-Guildwood",
+                "York South-Weston","Willowdale","Don Valley West",
+                "Don Valley North","Scarborough Centre")
+
+is_indigenous_garden_options <- c(0,1)
+
+# Create a dataset by randomly assigning values to variables for each id
+simulated_data <- tibble(
+  id = as.character(1:149),
+  year_funded = sample(
+    years_of_funding,
+    size = 149,
+    replace = TRUE
+  ),
+  garden_type = sample(
+    garden_types,
+    size = 149,
+    replace = TRUE
+  ),
+  number_of_gardens = sample(1:22, 149, replace = TRUE
+  ),
[…]

### 2024-12-04 scripts/01-test_simulated_data.R

@@ -1,0 +1,146 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the PollinateTO Primary Project Garden Locations simulated dataset
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+analysis_data_simulated <- read_csv("./data/00-simulated_data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data_simulated")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 149 rows
+if (nrow(analysis_data_simulated) == 149) {
+  message("Test Passed: The dataset has 149 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 149 rows.")
+}
+
+# Check if the dataset has 3 columns
+if (ncol(analysis_data_simulated) == 8) {
+  message("Test Passed: The dataset has 8 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 8 columns.")
+}
+
+# Check if all values in the 'id' column are unique
+if (n_distinct(analysis_data_simulated$id) == nrow(analysis_data_simulated)) {
+  message("Test Passed: All values in 'id' are unique.")
+} else {
+  stop("Test Failed: The 'id' column contains duplicate values.")
+}
+
+# Check if the 'garden_type' column contains only valid garden types
+valid_gardentype <- c("Multiple Front Yard Gardens",
+                    "Communal Garden",
+                    "School Learning Garden",
+                    "City Park",
+                    "Spiritual Centre Garden",
+                    "Boulevard Garden, Rain Garden",
+                    "Multi-Residential Garden",
+                    "Spiritual Centre Garden, Rain Garden",
+                    "Boulevard Garden",
+                    "City Park, Rain Garden",
+                    "Multiple Boulevard Gardens",
+                    "Multiple Property Garden")
+
+if (all(analysis_data_simulated$garden_type %in% valid_gardentype)) {
[…]

### 2024-12-04 scripts/02-download_data.R

@@ -1,0 +1,24 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data directly from Open Data Toronto
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+  # - The `tidyverse` and `opendatatoronto` packages must be installed and loaded
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+
+#### Download data ####
+package <- show_package("pollinateto-primary-project-garden-locations")
+resources <- list_package_resources("pollinateto-primary-project-garden-locations")
+csv_resource <- resources[resources$format == "CSV", ]
+first_csv_resource <- csv_resource[1, ]
+csv_id <- first_csv_resource$id
+pollinateto_data <- get_resource(csv_id) 
+
+#### Save data ####
+write_csv(pollinateto_data, "./data/01-raw_data/PT_raw_data.csv")

### 2024-12-04 scripts/03-clean_data.R

@@ -1,0 +1,55 @@
+#### Preamble ####
+# Purpose: Cleans the raw data directly from Open Data Toronto's website
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse`, `dplyr`, and `arrow` packages must be installed and loaded
+  # - 02-download_data.R must have been run
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+library(dplyr)
+library(arrow)
+
+#### Clean data ####
+PT_raw_data <- read_csv("./data/01-raw_data/PT_raw_data.csv")
+
+cleaned_data <-
+  PT_raw_data |>
+  janitor::clean_names() |>
+  # Remove the redundant 'x_id' column as it is same as 'id' column
+  select(-x_id) |>
+  # Remove columns that have no data i.e. all values are "None"
+  select (-description) |>
+  select (-image_name) |>
+  select (-image_alt_text) |>
+  # Remove unused columns
+  select (-group_name) |>
+  select (-project_name) |>
+  select (-neighbourhood_number) |>
+  select (-primary_location_name) |>
+  select (-primary_garden_address) |>
+  select (-postal_code) |>
+  select (-geometry) |>
+  select (-ward_number) |>
+  # ensure the year funded, number of gardens, estimated garden size are numbers
+  mutate(
+    year_funded = as.numeric(year_funded),
+    number_of_gardens = as.numeric(number_of_gardens),
+    estimated_garden_size = as.numeric(estimated_garden_size),
+    # ensure the nia_or_en, garden type, and ward name are strings
+    nia_or_en = as.character(nia_or_en),
+    garden_type = as.character(garden_type),
+    ward_name = as.character(ward_name),
+    # Convert the is_indigenous_garden column to binary column where 1 means that the project is indigenous led
+    is_indigenous_garden = if_else(is_indigenous_garden == "Y", 1, 0),
+    ward_name = if_else(ward_name == "Etobicoke-North", "Etobicoke North", ward_name)
+  ) |>
+  #drop any rows with missing/NA values
+  tidyr::drop_na()
+
+#### Save data ####
+write_parquet(cleaned_data, "./data/02-analysis_data/PT_analysis_data.parquet")

### 2024-12-04 scripts/04-test_analysis_data.R

@@ -1,0 +1,147 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the PollinateTO Primary Project Garden dataset
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+  # - The `tidyverse` and `arrow` packages must be installed and loaded
+  # - 02-download_data.R and 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(arrow)
+
+analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 149 rows
+if (nrow(analysis_data) == 149) {
+  message("Test Passed: The dataset has 149 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 149 rows.")
+}
+
+# Check if the dataset has 3 columns
+if (ncol(analysis_data) == 8) {
+  message("Test Passed: The dataset has 8 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 8 columns.")
+}
+
+# Check if all values in the 'id' column are unique
+if (n_distinct(analysis_data$id) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'id' are unique.")
+} else {
+  stop("Test Failed: The 'id' column contains duplicate values.")
+}
+
+# Check if the 'garden_type' column contains only valid garden types
+valid_gardentype <- c("Multiple Front Yard Gardens",
+                      "Communal Garden",
+                      "School Learning Garden",
+                      "City Park",
+                      "Spiritual Centre Garden",
+                      "Boulevard Garden, Rain Garden",
+                      "Multi-Residential Garden",
+                      "Spiritual Centre Garden, Rain Garden",
+                      "Boulevard Garden",
+                      "City Park, Rain Garden",
+                      "Multiple Boulevard Gardens",
+                      "Multiple Property Garden")
+
+if (all(analysis_data$garden_type %in% valid_gardentype)) {
[…]

### 2024-12-04 scripts/05-exploratory_data_analysis.R

@@ -1,0 +1,151 @@
+#### Preamble ####
+# Purpose: Examine trends, distributions, and relationships within the PollinateTO dataset
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+  # - The `tidyverse`, `arrow`, `ggplot2`, `knitr`, and `dplyr` packages must be installed and loaded
+  # - 02-download_data.R and 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+library(ggplot2)
+library(dplyr)
+library(knitr)
+library(arrow)
+
+#### Read data ####
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+# Summarize the data by year to get the total number of gardens and average garden size per year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded) %>%
+  summarise(
+    total_gardens = sum(number_of_gardens, na.rm = TRUE), # Summing number of gardens per year
+    avg_garden_size = mean(estimated_garden_size, na.rm = TRUE) # Calculating average garden size
+  )
+
+# Create the line chart with two y-axes (total gardens and average size)
+ggplot(PT_summary, aes(x = year_funded)) +
+  geom_line(aes(y = total_gardens), color = "blue", linewidth = 1.2, linetype = "solid") + 
+  geom_line(aes(y = avg_garden_size), color = "red", linewidth = 1.2, linetype = "dashed") +
+  scale_y_continuous(
+    name = "Total Number of Gardens", 
+    sec.axis = sec_axis(~., name = "Average Garden Size (Estimated)")
+  ) +
+  labs(
+    title = "Trend of Garden Size and Number Over Time",
+    x = "Year Funded",
+    y = "Total Number of Gardens"
+  ) +
+  theme_minimal() +
+  theme(
+    axis.title.y.right = element_text(color = "red"),
+    axis.text.y.right = element_text(color = "red")
+  )
+
+# Summarize the data to get the count of garden types by year
+PT_summary <- PT_analysis_data %>%
+  group_by(year_funded, garden_type) %>%
+  summarise(
[…]

### 2024-12-04 scripts/06-model_data.R

@@ -1,0 +1,49 @@
+#### Preamble ####
+# Purpose: Fits a Bayesian logistic regression model to predict whether a garden is large.
+# Author: Aliza Abbas Mithwani
+# Date: 3 December 2024
+# Contact: aliza.mithwani@mail.utoronto.ca
+# License: MIT
+# Pre-requisites:
+# - The `tidyverse`, `rstanarm`, and `arrow` packages must be installed and loaded
+# - 02-download_data.R and 03-clean_data.R must have been run
+# Any other information needed? Make sure you are in the `PollinateTO` rproj
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+library(arrow)
+
+# Load dataset
+PT_analysis_data <- read_parquet("./data/02-analysis_data/PT_analysis_data.parquet")
+
+
+# Prepare the data
+PT_analysis_data <- PT_analysis_data %>%
+  mutate(is_large_garden = ifelse(estimated_garden_size > mean(estimated_garden_size, na.rm = TRUE), 1, 0)) %>% # In binary form
+  mutate(
+    garden_type = as.factor(garden_type),
+    ward_name = as.factor(ward_name),
+    is_indigenous_garden = as.factor(is_indigenous_garden),
+    nia_or_en = as.factor(nia_or_en),
+    year_funded = as.numeric(year_funded) # Treated as continuous
+  )
+
+# Fit a Bayesian logistic regression model using rstanarm
+large_garden_model <- stan_glm(
+  is_large_garden ~ garden_type * nia_or_en + year_funded + is_indigenous_garden + ward_name, # Adding interaction between garden_type and nia_or_en
+  family = binomial(link = "logit"),
+  data = PT_analysis_data,
+  prior = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger normal priors
+  prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE), # Stronger prior on intercept
+  prior_aux = exponential(rate = 1),  # Regularizing the auxiliary parameters (error terms)
+  seed = 522,
+  chains = 4,  # Ensure more sampling chains
+  iter = 2000  # Set enough iterations for convergence
+)
+
+#### Save model ####
+saveRDS(
+  large_garden_model,
+  file = "./models/large_garden_model.rds"
+)

### 2024-12-15 other/datasheet/datasheet.qmd
Update LLM usage, Datasheet, References, Sketches

---
title: "Datasheet for 'PollinateTO Primary Project Garden Locations' Dataset"
subtitle: "An Overview of Garden Locations, Features, and Grant Information for PollinateTO Recipients"
author: 
  - Aliza Abbas Mithwani
thanks: "Code and data are available at: https://github.com/alizamithwani/PollinateTO.git."
date: today
date-format: "D MMMM YYYY"
abstract: "The PollinateTO Primary Project Garden Locations dataset documents the locations and characteristics of pollinator gardens funded through Toronto’s PollinateTO initiative. Analysis of this dataset reveals patterns in how garden funding is distributed across neighborhoods, including variations by ward, neighborhood type, and project characteristics. These findings highlight the potential influence of municipal funding priorities on urban biodiversity and community engagement. Understanding these patterns helps inform equitable urban greening efforts and supports the broader goal of fostering pollinator-friendly environments in cities."
format: pdf
number-sections: true
bibliography: references.bib
---

Extract of the questions from @gebru2021datasheets.
Questions answered using PollinateTO Primary Project Garden Locations dataset found on Open Data Toronto portal @PTdata

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to examine the allocation patterns of grants provided by the PollinateTO initiative in Toronto. The aim is to understand how various factors—such as pollinator types, neighborhood characteristics, and whether gardens are Indigenous-led—influence the size and distribution of community gardens. This analysis can inform future decision-making and equity considerations in grant allocations.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - The dataset is published by the City of Toronto's Environment & Climate department under the Open Government License - Toronto. The contact for the dataset is pollinateto@toronto.ca.
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - The funding for data creation likely stems from Toronto’s municipal budget for environmental and sustainability initiatives. Specific grant details are not provided but may relate to urban biodiversity programs.
4. *Any other comments?*
    - The dataset provides a valuable foundation for analyzing urban greening efforts and community-based sustainability practices.

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- Each instance represents a single PollinateTO-funded project, characterized by attributes such as location, grant size, garden size, urban density, and specific project details.
2. *How many instances are there in total (of each type, if appropriate)?*
	- The dataset contains 513 instances.
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- The dataset appears to represent the full set of PollinateTO-funded projects as of the date of collection. While it is unclear whether all relevant years and projects are included, the data spans multiple years and covers various wards and neighborhoods in Toronto. While the garden locations are publicly visible and accessible, some are on private property. This could affect analysis, particularly if any data points are not fully accessible for public engagement or observation.
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- Each instance consists of structured tabular data, including features such as the year, ward, pollinator type, grant size, garden size, urban density, and whether the garden is Indigenous-led.
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- There is no explicit “target” variable. However, fields such as “Grant size” or “Garden size” could serve as targets in predictive analyses.
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- There may be missing values in some columns (e.g., urban density or Indigenous-led status). Reasons for missing data are unclear but could include incomplete reporting or lack of data collection for certain fields.
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- No explicit relationships between instances are defined. However, some relationships may exist implicitly, such as multiple projects within the same ward or neighborhood.
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- No recommended data splits are provided. Data consumers could create splits based on temporal factors (e.g., by year) or geographic boundaries (e.g., wards).
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- The dataset appears clean at first glance. However, potential inconsistencies in categorical variables (e.g., “Pollinator type” or “Ward”) could introduce noise.
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- The dataset appears self-contained.
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- No. The data pertains to public grants and community gardens and does not include sensitive or confidential information.
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- No. The dataset is unlikely to contain any offensive or anxiety-inducing content.
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- The dataset identifies Indigenous-led gardens, which could be considered a sub-population. Distributions by ward and neighborhood are also available.
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- No. The data focuses on community gardens and grants rather than individual-level information.
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- No sensitive data is included. The only potentially sensitive field is whether a garden is Indigenous-led, but this is framed as a programmatic focus rather than an individual attribute.
16. *Any other comments?*
	- The dataset is straightforward and seems designed for public use and analysis.

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data appears to have been collected through administrative records from the PollinateTO initiative. These records likely document grant approvals and associated project details.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Data collection likely involved manual curation from grant applications and project reports submitted to the PollinateTO initiative, with potential use of spreadsheets or database software for recording project information.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The dataset does not appear to be a sample but rather includes all available PollinateTO projects up to the date of collection.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- Municipal staff, researchers, or grant administrators likely conducted data collection as part of their roles. Specific information about personnel and compensation is not provided.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The dataset spans multiple years, likely corresponding to the start of the PollinateTO initiative. The exact collection timeframe is unclear but seems aligned with the duration of the program.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- Ethical reviews were likely unnecessary, as the dataset pertains to public grants and does not include sensitive personal data.
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- The dataset does not directly relate to people.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- N/A as the dataset does not directly relate to people.
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- N/A as the dataset does not directly relate to people.
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- N/A as the dataset does not directly relate to people.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- N/A as the dataset does not directly relate to people.
12. *Any other comments?*
	- No other comments.

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- Minor preprocessing may have been conducted, such as standardizing categorical variables (e.g., ward names or pollinator types). However, the dataset primarily appears to be raw or minimally processed administrative data.
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- This is unclear. The dataset provided does not include any accompanying raw data.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- No preprocessing software is specified. If preprocessing was conducted, it may have involved common data tools such as Excel or R.
4. *Any other comments?*
	- No other comments.

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- The dataset has not yet been used for published analyses but is intended for examining factors influencing grant allocations and the distribution of PollinateTO-funded gardens.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- No known repository exists at this time.
3. *What (other) tasks could the dataset be used for?*
	- Since the dataset is openly available under a public license, it can be freely used for analysis, but any use of private property data might need to be considered in the analysis to ensure compliance with privacy expectations. The dataset could be used to: Analyze spatial patterns of urban greening efforts; Evaluate equity in grant distribution across neighborhoods; Identify trends in funding for Indigenous-led projects; Explore the impact of urban density on garden sizes and types.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- Consumers should be cautious when interpreting variables like "Indigenous-led" gardens to avoid unintended implications about communities or funding fairness. Additionally, variations in reporting quality across wards could bias results.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- The dataset should not be used to draw conclusions about individual-level behavior or to assess the personal characteristics of those involved in projects.
6. *Any other comments?*
	- No other comments

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- Yes, the dataset is intended for public analysis and research purposes.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset could be shared via public repositories (e.g., GitHub) or municipal open data portals. No DOI is currently available.
3. *When will the dataset be distributed?*
	- The distribution timeline is unspecified, but it is likely available immediately or soon after compilation.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- The dataset is published under the Open Government License - Toronto. The Open Government License - Toronto allows users to freely use, modify, and share datasets published by the City of Toronto, with the condition that the source is properly credited. The license permits both commercial and non-commercial uses, but disclaims any warranties regarding the accuracy or completeness of the data, and prohibits implying endorsement by the City of Toronto. There are no fees associated with the license, and users are free to create derivative works, as long as they meet the attribution requirement. Full details of the license can be found on the City of Toronto's Open Data Portal.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No IP-based restrictions are indicated.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No export or regulatory restrictions apply.
7. *Any other comments?*
	- No other comments.

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- Likely the municipal government of Toronto or affiliated researchers.
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- Contact information is not provided but could potentially be found via municipal or research channels.
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No errata are available at this time.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- The dataset could be updated annually to include new PollinateTO projects. Updates might be communicated via public portals or repositories. The dataset was last refreshed on October 15, 2024, and is updated annually.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- The dataset does not relate to individuals, so retention limits are not applicable.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- It is unclear if older versions will be maintained, though retaining historical records could aid longitudinal studies.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- Collaboration mechanisms are not mentioned but could involve municipal partnerships or public data contribution tools.
8. *Any other comments?*
	- The dataset offers valuable opportunities for research into urban sustainability practices and grant distribution equity.

\newpage

# References

### 2024-12-15 paper/paper.qmd
Update paper

@@
 ---
-title: "Key Drivers of Community Garden Growth: Analyzing Trends in PollinateTO Funding"
+title: "Factors Influencing Community Garden Growth: Analyzing Trends in PollinateTO Funding"
 subtitle: "A Bayesian Approach to Understanding the Impact of Time, Type, and Location on Garden Size and Number in Toronto's Neighborhoods from 2020 to 2023."
@@
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "This paper investigates the allocation patterns of grants for pollinator gardens funded by the PollinateTO initiative in Toronto, focusing on how garden characteristics, neighborhood type, and community engagement influence funding decisions.
+A Bayesian logistic regression model reveals that gardens located in Neighbourhood Improvement Areas (NIA) or Emerging Neighborhoods (EN) are more likely to receive higher funding, especially for specific garden types like Multi-Residential Gardens and Multiple Property Gardens.
+Additionally, Indigenous-led gardens and those located in certain wards (e.g., Scarborough Centre) tend to receive larger grants.
+This research contributes to optimizing the allocation of urban biodiversity funding, ensuring equitable support for community-driven ecological projects."
 format:
@@
 library(patchwork)
-library(tinytable)
 library(kableExtra)
@@
 
-## Overview
-Urban biodiversity is increasingly recognized as vital to maintaining healthy ecosystems, especially in the face of rapid urbanization.
-Among the most important urban biodiversity goals is the protection of pollinators, whose decline poses significant risks to food security, agricultural productivity, and ecological stability.
-In response to this, many cities have launched initiatives to create pollinator-friendly habitats within urban spaces.
+Urban biodiversity is recognized as vital to maintaining healthy ecosystems, particularly in the context of rapid urbanization.
+A key goal is the protection of pollinators, whose decline presents risks to food security, agricultural productivity, and ecological stability.
+In response to this, many cities have launched initiatives to create pollinator-friendly habitats within urban areas.
 Toronto’s PollinateTO program is one such initiative, aiming to enhance biodiversity by funding community-led pollinator gardens.
@@
 
-Despite the program's success in expanding green spaces, the allocation of funding across these projects remains an understudied area.
-Questions about which types of gardens or neighborhoods are more likely to receive funding, and the factors influencing these decisions, are crucial for optimizing the program’s impact.
+Despite the program's success in expanding green spaces, the allocation of funding across these projects remains an under-explored area.
+Questions about which types of gardens or neighborhoods are more likely to receive funding, and the factors influencing these decisions, are important for optimizing the program’s impact.
 The objective of this paper is to examine these allocation patterns, specifically whether garden characteristics such as type (e.g., rain gardens, boulevard gardens) or location (e.g., Neighbourhood Improvement Areas) influence the likelihood of receiving funding and the size of the funding allocation.
 
-In particular, this paper addresses the gap in understanding how urban biodiversity funding decisions are made, with a focus on the specific factors that contribute to higher funding likelihoods and amounts.
-The estimand in this analysis is whether particular garden features, community engagement strategies, or geographic locations are associated with receiving more substantial funding allocations from the PollinateTO program.
-By identifying these key factors, this research provides important insights into the effectiveness and equity of funding distributions for urban biodiversity projects.
+This paper addresses the lack of understanding how urban biodiversity funding decisions are made, with a focus on the factors that contribute to higher funding likelihoods and amounts.
+The estimand in this analysis is whether particular garden features, community engagement strategies, or geographic locations are associated with receiving larger funding allocations from the PollinateTO program.
+By identifying these factors, this research provides important findings related to the effectiveness and equity of funding distributions for urban biodiversity projects.
 
 Our findings indicate that garden type, community engagement, and the location of the project all play significant roles in funding decisions.
-Specifically, we find that gardens in underserved neighborhoods and those emphasizing community involvement are more likely to receive higher funding allocations.
-These results offer actionable insights for optimizing resource distribution in urban biodiversity programs and aligning ecological goals with social equity considerations.
+Specifically, we find that gardens in underserved neighborhoods and those emphasizing community involvement are more likely to receive larger funding allocations.
+These results offer actionable recommendations for optimizing resource distribution in urban biodiversity programs and aligning ecological goals with social equity considerations.
 
 The remainder of the paper is structured as follows: @sec-data describes the data used in the analysis, including its sources, key features, and considerations regarding measurement and visualization.
-Section 3 outlines the methodology, detailing the data visualization techniques employed and the justification for using a Bayesian logistic regression model to analyze the relationship between garden characteristics and funding outcomes.
-Section 4 presents the results, including model validation and key findings.
-Section 5 offers a discussion of the results, focusing on their implications through key points, followed by an examination of the study's limitations and suggested future steps for further exploration.
-Section 6 concludes the paper.
-The Appendix provides supplementary materials, including raw data previews, data feature definitions, data visualizations, and technical model details such as posterior predictive checks, MCMC convergence diagnostics, and additional figures.
+@sec-model outlines the methodology, detailing the data visualization techniques employed and the justification for using a Bayesian logistic regression model to analyze the relationship between garden characteristics and funding outcomes.
+@sec-results presents the results, including model validation and key findings.
+@sec-discussion offers a discussion of the results, focusing on their implications through key points, followed by an examination of the study's limitations and suggested future steps for further analysis.
+@sec-finals concludes the paper.
+The {-} provides supplementary materials, including additional model details related to cleaning and simulation, a section about survey, sampling, and observational data in the context of this paper, and technical model details such as posterior predictive checks and MCMC convergence diagnostics.
 
@@
 ## Software Packages
-This analysis utilizes the R programming language [@citeR], with the `rstanarm` [@rstanarm] and `tidyverse` [@tidyverse] packages for statistical modeling and data manipulation.
+This analysis utilizes the R programming language [@citeR], with the `rstanarm` [@rstanarm], `tidyverse` [@tidyverse], `opendatatoronto` [@opendatatoronto], `ggplot2` [@ggplot2], `dplyr` [@dplyr], `knitr` [@knitr], `arrow` [@arrow], `patchwork` [@patchwork], `kableExtra` [@kableExtra], and `broom.mixed` [@broommixed] packages for statistical modeling and data manipulation.
 Our data comes from the PollinateTO Primary Project Garden Locations dataset [@PTdata], which is hosted on Toronto's Open Data Portal.
-Following the approach outlined in @tellingstories, we leverage these tools to explore patterns and trends in the allocation of PollinateTO funding.
+Following the approach outlined in @tellingstories, we use these tools to explore patterns and trends in the allocation of PollinateTO funding.
 
 ## Data Source
-The primary source of data for this study is the PollinateTO Primary Project Garden Locations dataset, which contains detailed information on community-led garden projects funded by PollinateTO.
+The primary source of data for this study is the PollinateTO Primary Project Garden Locations dataset [@PTdata], which contains detailed information on community-led garden projects funded by PollinateTO.
 The dataset includes key features such as project locations, garden types, funding years, and other relevant information that enables an in-depth analysis of funding patterns and their spatial and socio-economic implications.
@@
 ## Data Features
-The dataset contains several key variables that are critical for understanding the distribution and characteristics of funded garden projects.
+The dataset contains several key variables that are important for understanding the distribution and characteristics of funded garden projects.
 Below, we describe each variable and its relevance to the analysis:
@@
 - YEAR_FUNDED: The year each project received funding from PollinateTO.
-This variable is crucial for understanding temporal trends in the program’s funding allocation.
+This variable is important for understanding temporal trends in the program’s funding allocation.
 It also allows us to analyze shifts in funding priorities over time.
@@
 - NIA_OR_EN: This classification indicates whether the project is located in a Neighbourhood Improvement Area (NIA), Emerging Neighbourhood (EN), or another area.
-These labels are derived from Toronto's planning data and allow for an exploration of whether funding is disproportionately allocated to areas with different socio-economic characteristics, enabling us to analyze equity in the distribution of funds.
+These labels are derived from Toronto's planning data and allow for a search of whether funding is disproportionately allocated to areas with different socio-economic characteristics, enabling us to analyze equity in the distribution of funds.
 - WARD_NAME: The specific ward in Toronto where the garden project is located.
@@
 This is a key variable for exploring whether the program is aligned with Toronto’s commitments to Indigenous communities and ecological knowledge.
-PROJECT_NAME: The name assigned to each garden project.
+- PROJECT_NAME: The name assigned to each garden project.
 While this variable is useful for identification, it is not directly involved in the analysis but helps in linking data points to specific projects.
@@
 ## Constructed Variables
-In this study, variable is_large_garden was constructed.
-This binary variable was constructed to classify gardens as "large" or "small" based on their estimated garden size.
+In this study, the variable is_large_garden was constructed.
+This binary variable was created to classify gardens as "large" or "small" based on their estimated garden size.
 Specifically, gardens with an estimated garden size greater than the mean size of all gardens in the dataset were coded as 1 (large garden), while those with a size less than or equal to the mean were coded as 0 (small garden).
@@
 Granularity: Some ecological measures, such as specific plant species or detailed environmental outcomes, are not captured in the dataset.
-This restricts the scope of ecological impact analysis, though the dataset’s focus on garden characteristics and funding allocation provides valuable insights.
+This restricts the scope of ecological impact analysis, though the dataset’s focus on garden characteristics and funding allocation provides important information.
 
@@
 Data is sourced from the city’s Open Data portal and includes variables like garden type, location, funding year, and size, enabling analysis of trends in urban biodiversity projects.
-The dataset uses standardized attributes and is updated annually, making it a valuable tool for evaluating the initiative’s impact across various Toronto neighborhoods, though it may be subject to self-reporting biases and geographical limitations in precision.
+The dataset uses standardized attributes and is updated annually, making it a useful tool for evaluating the initiative’s impact across various Toronto neighborhoods, though it may be subject to self-reporting biases and geographical limitations in precision.
 \newpage
@@
 ```
-# Model
+# Model {#sec-model}
 In the analysis, I used a Bayesian Logistic Regression Model to examine the relationship between garden size and other factors like the type of garden, the year the project was funded, the neighborhood type (emerging neighborhood, EN vs neighborhood improvement area, NIA), and if it is an indigenous-led project.
 We are interested in modeling whether a garden is large (binary outcome) as a function of various factors such as the garden type, year funded, neighborhood type, indigenous garden status, and ward name.
-Background details and diagnostics are included in [Appendix -@sec-model-details].
+Background details and diagnostics are included in [Appendix - @sec-model-details].
 
@@
 \beta_4 &\sim \mbox{Normal}(0, 2.5) \\
-\beta_5 &\sim \mbox{Normal}(0, 2.5)
+\beta_5 &\sim \mbox{Normal}(0, 2.5) \\
+\beta_6 &\sim \mbox{Normal}(0, 2.5)
 \end{align}
@@
 
-### Model justification
+## Model justification
 
@@
 
-The choice of the Bayesian approach in this context also allows for the flexible inclusion of multiple categorical predictors, which is crucial for this study, as it aims to assess the relationship between categorical factors like garden type, ward names, and the interaction between garden type and nia or en with a quantitative outcome.
+The choice of the Bayesian approach in this context also allows for the flexible inclusion of multiple categorical predictors, which is important for this study, as it aims to assess the relationship between categorical factors like garden type, ward names, and the interaction between garden type and nia or en with a quantitative outcome.
 The model's use of informative priors helps regularize the estimates, particularly when dealing with categorical predictors that may have complex relationships with the outcome variable.
-Additionally, the Bayesian methodology provides the opportunity to assess model diagnostics comprehensively, ensuring that the model fits well with the data and that the parameters are estimated with high precision.
+Additionally, the Bayesian methodology provides the opportunity to assess model diagnostics, ensuring that the model fits well with the data and that the parameters are estimated with high precision.
 Overall, this model is well-suited to capture the complexities of the data while offering interpretability and flexibility through its probabilistic nature.
 
-# Results
+# Results {#sec-results}
 
 Our results are summarized in @tbl-modeltable.
-The Bayesian logistic regression model reveals several important relationships between garden characteristics and the likelihood of a garden being classified as large.
+The Bayesian logistic regression model shows several important relationships between garden characteristics and the likelihood of a garden being classified as large.
 Among garden types, "Multiple Property Garden" stands out with a substantial positive coefficient (14.94), suggesting that such gardens are strongly associated with larger sizes.
@@
 #| warning: false
-#| tbl-cap: "Explanatory model of whether a garden is large based on certain factors: Indigenous-led indicator, Year Funded, Neighborhood, Garden type and Ward Names"
+#| tbl-cap: "Explanatory model of whether a garden is large based on certain factors: Indigenous-led indicator, Year Funded, Neighborhood, Garden type and Ward Names (zoom in for a clearer view)"
 
@@
   kable_styling(latex_options = c("scale_down", "hold_position")) %>%
-  landscape()
+  landscape() %>%
+  column_spec(1, width = "1cm") %>%
+  kable_styling(font_size = 3)  # Reduce the font size to make it smaller
 ```
@@
 Predictors whose credible intervals do not cross the dashed red line at zero are likely to have significant effects.
-This plot highlights which variables, such as garden type, neighborhood classification, or year funded, are most influential in determining garden size, providing valuable insights into the factors driving outcomes in the program.
+This plot highlights which variables, such as garden type, neighborhood classification, or year funded, are most influential in determining garden size, providing understanding of the factors driving outcomes in the program.
 
@@
 #### Predictive Probability Visualization ####
-
-# Load necessary libraries
-library(tidyverse)
-library(ggplot2)
-library(rstanarm)
 
@@
 \newpage
-# Discussion
+# Discussion {#sec-discussion}
 
 ## Unearthing Patterns in Garden Allocation {#sec-first-point}
+
 This paper investigates the factors influencing the size of gardens funded under Toronto’s PollinateTO initiative using a Bayesian logistic regression model.
-Specifically, it explores how garden size (a binary outcome of “large” vs.
-“not large”) is associated with variables such as garden type, year funded, neighborhood classification (Emerging Neighborhoods or Neighborhood Improvement Areas), Indigenous-led status, and ward-level geographic differences.
-By leveraging this model, the analysis identifies patterns and disparities in garden allocation, providing insights into which factors are most predictive of larger gardens.
-The approach incorporates interaction effects to capture nuanced relationships between garden types and neighborhood characteristics.
+Specifically, it explores how garden size, defined as a binary outcome (large vs.
+not large), is associated with several key predictors: garden type, year funded, neighborhood classification (Emerging Neighborhoods or Neighborhood Improvement Areas), Indigenous-led status, and ward-level geographic differences.
+The analysis shows several noteworthy patterns and disparities in the allocation of larger gardens.
+By leveraging this model, the study identifies which factors have the most significant influence on whether a garden is classified as large.
+This approach is enhanced by incorporating interaction effects, which provide a deeper understanding of the relationships between garden types and neighborhood characteristics.
 
-The inclusion of credible intervals from the Bayesian model ensures a robust interpretation of uncertainty in coefficient estimates.
-Additionally, key goodness-of-fit measures (e.g., R², WAIC) are evaluated, demonstrating the model’s ability to explain nearly half of the variation in garden size.
-These findings offer a data-driven lens through which urban sustainability programs like PollinateTO can be assessed and improved.
+The Bayesian logistic regression framework offers a probabilistic interpretation of these relationships, allowing for a more complex view of uncertainty in parameter estimates.
+By incorporating credible intervals, the model quantifies the uncertainty around the estimated coefficients, providing a more robust and transparent analysis compared to traditional frequentist methods.
+In addition to providing understanding of the predictive factors of garden size, this paper also assesses model fit through key measures such as R² and WAIC.
+These measures indicate that the model explains a significant portion of the variation in garden size, with nearly half of the variation accounted for by the chosen predictors.
+
+The implications of these findings are significant for urban sustainability programs like PollinateTO.
+By highlighting the key drivers of garden size, this research offers an analytical lens through which policymakers can refine the allocation process for funding community gardens.
+If the goal of such initiatives is to foster environmental sustainability and equity, then understanding which factors contribute to larger gardens is important.
+For instance, urban greening efforts that are situated in more resource-rich neighborhoods or that prioritize certain types of gardens may need to be adjusted to address inequities and ensure that all communities have access to the same opportunities for greening and sustainability.
+
+Additionally, the interaction terms between garden type and neighborhood classification are particularly important.
+They show that certain garden types (e.g., vegetable or pollinator gardens) tend to be more successful in certain types of neighborhoods, which may point to localized resources or specific neighborhood priorities.
+This finding suggests that urban greening programs should consider tailoring their strategies based on local needs and resources.
+Programs that adopt a more context-sensitive approach, considering both neighborhood characteristics and garden types, are likely to see more equitable and effective outcomes.
 
@@
 
-One of the central insights from this analysis is the existence of disparities in garden size across neighborhoods.
-Specifically, gardens in areas classified as “None” (neither EN nor NIA) are significantly more likely to be large, while gardens in NIAs face challenges in achieving similar sizes.
-This suggests that systemic factors, such as socioeconomic inequalities, land availability, and access to resources, influence the outcomes of urban greening initiatives.
+A major finding from this analysis is the evident disparities in garden size across different neighborhoods.
+Specifically, gardens in areas classified as “None” (i.e., neighborhoods neither in the Emerging Neighborhoods nor in the Neighborhood Improvement Areas) are significantly more likely to be classified as large, while gardens located in Neighborhood Improvement Areas (NIAs) are less likely to achieve the same size.
+This finding suggests that neighborhoods with fewer systemic challenges, such as better access to resources, land availability, and socioeconomic stability, are more likely to develop large, successful urban gardens.
+Conversely, communities in NIAs, which often face challenges like poverty, limited access to green space, and lower levels of community investment, struggle to achieve similar outcomes.
 
-This insight highlights a broader inequity in urban sustainability programs: neighborhoods with fewer resources or systemic disadvantages may struggle to achieve comparable environmental benefits.
-This finding emphasizes the importance of embedding equity-focused considerations into grant allocation processes.
-Addressing these disparities can improve not only the environmental impact of such initiatives but also their social relevance and inclusivity.
+This disparity underscores a broader issue in urban sustainability programs: systemic factors, such as socioeconomic inequalities, land availability, and access to resources, significantly affect the success of these initiatives.
+These inequalities are often compounded by historical disinvestment in marginalized neighborhoods, which can lead to lower levels of community engagement and reduced capacity for sustaining urban greening efforts.
+The difference in garden size between these neighborhoods points to the need for equity-focused policies in urban greening initiatives.
+
+The implication is clear: sustainability programs must address neighborhood-level disparities in access to resources and opportunities.
+While it is important to continue fostering urban sustainability, it is equally important to ensure that these efforts benefit all communities, particularly those that have historically been underserved.
+Programs like PollinateTO should consider revising their funding allocation strategies to prioritize neighborhoods with fewer resources or greater systemic disadvantages.
+This could involve providing additional technical assistance, offering capacity-building programs for community members, or targeting more funding toward NIAs to ensure these communities can achieve similar environmental and social benefits.
+By doing so, urban sustainability programs can more effectively address environmental justice concerns, ensuring that all communities have the opportunity to benefit from urban greening.
+
+Moreover, this finding points to a need for broader systemic change.
+To truly address inequities in urban gardening, policymakers must also focus on improving the underlying conditions in marginalized neighborhoods.
+This could include initiatives to improve access to affordable housing, better public transportation, and community infrastructure.
+Without such systemic changes, urban greening programs may fail to achieve their full potential in creating equitable and sustainable urban environments.
 
@@
 
-Indigenous-led projects are less likely to result in large gardens, pointing to potential barriers that Indigenous communities face in participating fully in urban greening initiatives.
-This finding underscores the need to reevaluate the criteria for “success” in such programs.
-Indigenous-led projects may prioritize cultural and ecological goals—such as the preservation of native plant species or community education—over garden size.
+Indigenous-led projects in urban greening initiatives, including those funded by PollinateTO, are less likely to result in large gardens compared to other projects.
+This finding raises important questions about the criteria used to evaluate the success of these projects.
+Indigenous communities may prioritize different values in their urban greening efforts, focusing on goals such as the preservation of native plant species, the restoration of traditional ecological practices, and the fostering of community cohesion, rather than on maximizing garden size.
+As a result, the existing framework for evaluating the success of urban gardens, which focuses primarily on the quantitative measure of size, may not fully capture the diverse goals and impacts of Indigenous-led initiatives.
 
-This raises an important consideration for future urban sustainability programs: qualitative impacts, such as community engagement or cultural significance, may be equally (if not more) important than quantitative measures like garden size.
-By failing to account for these dimensions, existing frameworks risk marginalizing projects that offer critical social and cultural benefits.
-This insight encourages a shift in how success is defined and measured in urban greening efforts.
+This finding calls for a reevaluation of how success is defined and measured in urban sustainability programs.
+While garden size is a useful indicator of certain types of success, it may not adequately capture the full range of outcomes that Indigenous-led projects aim to achieve.
+For example, many Indigenous communities prioritize cultural, social, and ecological sustainability over the mere expansion of garden size.
+Gardens that are smaller but culturally significant or that provide vital ecosystem services may be just as important, if not more so, than larger gardens that lack these dimensions.
 
-# Weaknesses and next steps
-I will address some of the weaknesses and limitations of the current analysis, focusing on areas where the methodology or data could be improved.
-Following this, I will explore potential next steps, outlining opportunities for future research and policy development that build on these findings to create more equitable and effective urban sustainability initiatives.
-This section aims to critically reflect on the study while charting a path forward for both academic inquiry and practical applications.
+The need to redefine success also speaks to a broader issue in urban sustainability programs: the tendency to prioritize Western, quantitative metrics of success, such as garden size, over qualitative outcomes that may be more meaningful to Indigenous communities.
+This discrepancy often leads to the marginalization of projects that offer important cultural and social benefits but do not conform to conventional measures of success.
+By failing to account for these dimensions, urban sustainability frameworks risk overlooking initiatives that are culturally relevant and community-driven.
+
+This finding suggests a shift toward a more inclusive and holistic approach to evaluating urban greening projects.
+Success should not be measured solely by garden size, but rather by the broader social, cultural, and environmental impacts that these gardens have on the community.
+Future iterations of urban greening programs should consider the inclusion of qualitative measures, such as community engagement, cultural preservation, and ecological restoration, alongside traditional quantitative metrics.
+This would ensure that Indigenous-led projects are evaluated on their own terms and that they receive the recognition and support they deserve.
+
+# Weaknesses and next steps {#sec-finals}
+
+While this analysis provides important findings, it is important to acknowledge the limitations of the current study and suggest areas for improvement in both the methodology and data collection.
+Addressing these weaknesses will help refine the findings and lead to more robust and equitable urban sustainability programs.
 
 ## Challenges and Limitations in Modeling Garden Outcomes
-While the model provides valuable insights, it is not without limitations.
-First, the binary categorization of garden size as “large” vs.
-“not large” may oversimplify the complexity of garden outcomes.
-A more nuanced approach, such as modeling size as a continuous variable or including qualitative measures of success, might provide richer insights.
 
-Second, the dataset is limited to gardens funded by PollinateTO, which may not fully represent the broader urban gardening landscape in Toronto.
-Selection bias could influence the generalizability of these findings to other urban greening initiatives.
-Additionally, the potential for omitted variable bias exists, as factors like community engagement levels or prior gardening experience were not included in the analysis but may significantly impact garden outcomes.
+The model employed in this study provides a useful understanding of the factors influencing garden size, but it has several limitations.
+One of the primary weaknesses is the binary categorization of garden size into “large” vs.
+“not large.” This binary approach oversimplifies the complexity of garden outcomes, as garden size is a continuous variable that may better reflect the true diversity of urban gardens.
+Future research could model garden size as a continuous variable, allowing for a deeper understanding of how factors such as neighborhood characteristics and garden type influence garden size.
+Additionally, including qualitative measures of success, such as the impact of the garden on local biodiversity or community engagement, would enrich the analysis and provide a more thorough understanding of garden outcomes.
 
-Finally, while interaction terms capture some complexity, they add interpretive challenges, as not all combinations of garden types and neighborhoods had sufficient data to provide precise estimates.
-Future work could address this by gathering larger, more diverse datasets or employing hierarchical models to account for varying group sizes.
+Another limitation is the focus on gardens funded by the PollinateTO initiative.
+While this dataset provides important findings, it may not fully represent the broader urban gardening landscape in Toronto.
+The sample is limited to one specific initiative, and the findings may not be generalizable to other urban greening programs or gardening projects outside of Toronto.
+Moreover, the potential for selection bias exists, as the gardens included in the dataset are those that were selected for funding, and these may not be representative of all gardens in the city.
+Future studies could broaden the scope to include data from other urban greening initiatives or include non-funded gardens to improve the generalizability of the findings.
+
+Finally, while interaction terms in the model provide some understanding of how garden types and neighborhood characteristics interact, they add interpretive complexity.
+Not all combinations of garden types and neighborhoods had sufficient data to provide precise estimates.
+In particular, some combinations of garden type and neighborhood classification may have limited representation, leading to imprecise estimates.
+Future research could address this by gathering larger and more diverse datasets or employing hierarchical models that account for varying group sizes and complexities.
 
 ## Opportunities for Future Research and Policy Improvements
-Several open questions remain.
-One key area for future research is the inclusion of community-level variables, such as demographic characteristics, access to volunteer networks, and proximity to green spaces.
-These factors may help explain why certain neighborhoods are more likely to host large gardens.
+Several open questions remain that could be explored in future research.
+One key area is the inclusion of community-level variables, such as demographic characteristics, access to volunteer networks, and proximity to green spaces.
+These factors could help explain why certain neighborhoods are more likely to host large gardens and may show underlying mechanisms driving these disparities.
+Incorporating demographic data could also help policymakers better understand the unique challenges and opportunities faced by different communities in relation to urban greening.
 
-Additionally, further work is needed to explore the qualitative outcomes of urban gardens, particularly for Indigenous-led projects.
-Incorporating interviews, case studies, or community feedback could provide a deeper understanding of how these gardens contribute to cultural preservation and community well-being.
+In addition to these community-level factors, further research should focus on exploring the qualitative outcomes of urban gardens, particularly for Indigenous-led projects.
+Incorporating interviews, case studies, or community feedback could provide a richer understanding of how these gardens contribute to cultural preservation, social cohesion, and community well-being.
+By capturing the voices of community members, future studies can gain a more holistic understanding of the impacts of urban greening initiatives, particularly in Indigenous communities.
 
-From a methodological perspective, future studies could explore alternative modeling techniques, such as hierarchical Bayesian models, to better account for ward-level variability and cross-neighborhood interactions.
-Expanding the dataset to include other urban greening initiatives or longitudinal data could also provide insights into how garden outcomes evolve over time.
+From a methodological lens, future research could explore alternative modeling techniques, such as hierarchical Bayesian models, to better account for ward-level variability and cross-neighborhood interactions.
+This would allow for deeper interpretations of how different neighborhoods and wards contribute to garden outcomes.
+Expanding the dataset to include data from other urban greening initiatives or longitudinal data that tracks changes over time would also provide important understanding of how garden outcomes evolve and whether the impact of different factors changes over time.
 
-Lastly, there is a need to test the effectiveness of policy interventions informed by these findings.
-For instance, targeted funding for NIAs or revised evaluation criteria for Indigenous-led projects could be piloted and evaluated to determine their impact on garden size and community engagement.
+Finally, there is a need for future studies to test the effectiveness of policy interventions informed by these findings.
+For example, targeted funding for Neighborhood Improvement Areas (NIAs) or revised evaluation criteria for Indigenous-led projects could be piloted and evaluated to determine their impact on garden size, community engagement, and other outcomes.
 This iterative process of evidence-based policy design and evaluation can help create more equitable and effective urban sustainability programs.
+By testing different policy interventions, cities can identify the most effective strategies for promoting urban greening and achieving more equitable outcomes for all communities.
 
@@
 
+The cleaning process for the PollinateTO dataset involved several key steps to ensure the data was prepared for analysis.
+Initially, the raw dataset, downloaded from Open Data Toronto, was read into R.
+The janitor package was used to clean column names for consistency and readability.
+Redundant columns, such as 'x_id', which duplicated the 'id' column, were removed to streamline the dataset.
+Additional irrelevant columns, including 'description', 'image_name', 'postal_code', and others containing only missing values or unnecessary information, were also discarded.
+Columns such as 'year_funded', 'number_of_gardens', and 'estimated_garden_size' were explicitly converted to numeric formats to ensure proper data types, while 'nia_or_en', 'garden_type', and 'ward_name' were converted to character strings.
+The 'is_indigenous_garden' column was recoded as a binary variable, with 1 indicating Indigenous-led projects and 0 otherwise.
+Furthermore, any rows with missing values were removed to maintain the integrity of the dataset.
+Lastly, the cleaned data was saved in Parquet format for efficient storage and further analysis.
+
 ## Data Simulation
+
+To simulate the dataset for this study, a random sampling approach was employed to generate synthetic data that mirrors the characteristics of the PollinateTO initiative's garden locations.
+The simulation process involved assigning values to several key variables, including the year of funding, garden type, the number of gardens, estimated garden size, neighborhood classification (NIA, EN, or None), Indigenous-led status, and ward name.
+For each of these variables, values were randomly selected from predefined lists.
+For example, the "year_funded" variable was drawn from the years 2020 to 2023, while garden types were chosen from a list that includes a variety of garden types such as "Communal Garden" and "Spiritual Centre Garden." Additionally, random values were generated for the number of gardens and estimated garden sizes within realistic ranges.
+To ensure variability across different wards and neighborhood classifications, the "nia_or_en" and "ward_name" variables were sampled from their respective options.
+The Indigenous-led status was simulated as a binary variable with values of 0 or 1.
+The resulting dataset includes 149 entries, with each row representing a simulated garden project.
+The simulated data was then saved for further analysis, ensuring it captures the diversity and distribution of factors relevant to the PollinateTO initiative.
+
+# Surveys, Sampling, and Observational Data
+In this section, we go into the methodology underlying the dataset used for this study, which focuses on analyzing the factors influencing the size and allocation of gardens funded by the City of Toronto's PollinateTO initiative.
+The data, derived from the open-source dataset on primary project garden locations provided by the City of Toronto, offers a unique opportunity to examine the relationships between various community and garden-level characteristics.
+This dataset includes information on garden size, neighborhood classifications, garden types, and the presence of Indigenous-led gardens, which serves as the foundation for the analysis in this paper.
+However, the observational nature of the data raises important questions about sampling methods, biases, and generalizability, which we address below.
+
+## Survey and Data Collection Methodology
+The PollinateTO dataset is primarily observational, as it includes data from gardens that were funded through the initiative.
+However, while the dataset provides a rich source of information, there are several aspects of the data collection process that must be scrutinized.
+First, the data is not drawn from a random sample of all urban gardens in Toronto, but specifically from those that received funding from the PollinateTO initiative.
+This introduces a potential selection bias in the dataset, as gardens that are funded by PollinateTO may differ from those that are not, in terms of size, location, and community characteristics.
+The lack of randomization in the selection of gardens funded by the initiative means that any conclusions drawn from the data may be limited to this specific subset of gardens, potentially reducing the external validity of the findings when applied to broader urban gardening contexts.
+
+The dataset includes key variables such as garden type, neighborhood classification (Emerging Neighborhoods or Neighborhood Improvement Areas), the year funding was provided, and the ward in which each garden is located.
+These variables were likely chosen to capture factors that influence garden success and size, but it is important to note that other relevant variables, such as community engagement or access to resources, are not included.
+The absence of these factors may lead to omitted variable bias, as unmeasured factors could also influence garden size, and therefore the results of this study may only be partial.
+
+## Sampling Considerations and Biases
+In terms of sampling, the PollinateTO initiative targets gardens that aim to support pollinator species and contribute to biodiversity in urban areas.
+The initiative is aligned with the broader goals of sustainability and environmental resilience in Toronto, and it emphasizes inclusivity by supporting gardens in various types of neighborhoods.
+However, the program's funding decisions are likely influenced by factors such as available funding, the perceived urgency of greening initiatives in specific neighborhoods, and the organizational capacity of communities to implement and maintain gardens.
+As a result, neighborhood characteristics such as socioeconomic status, access to green spaces, and local community engagement in sustainability initiatives may systematically influence which areas receive funding, potentially skewing the data towards neighborhoods with higher resources or stronger community organizations.
+
+Furthermore, because the dataset is observational rather than experimental, the model used in this study must account for potential confounding variables.
+For instance, neighborhood-level characteristics like income inequality or infrastructure quality may be correlated with both garden size and the likelihood of a neighborhood receiving PollinateTO funding, yet these factors are not explicitly included in the dataset.
+This type of confounding bias can make it challenging to draw causal inferences from the analysis, as the relationships between the predictors (e.g., garden type, neighborhood type, and ward location) and garden size may be influenced by other, unobserved factors.
+
+## Methodological Considerations and Simulation
+To address some of these biases and limitations, future research could incorporate simulation techniques to generate counterfactual scenarios, such as estimating the outcomes for neighborhoods that did not receive PollinateTO funding.
+This approach would help assess the extent of selection bias and better understand how the distribution of gardens might have differed under alternative conditions.
+Moreover, propensity score matching could be employed to create a matched sample of funded and non-funded gardens, ensuring that both groups are comparable on observed characteristics, thereby mitigating potential confounding effects.
+
+Additionally, longitudinal data on garden outcomes, including follow-up data on garden size and sustainability, would provide more understanding of how factors like neighborhood characteristics and garden types evolve over time.
+This would help disentangle short-term effects of funding from long-term impacts, and improve the understanding of how urban sustainability programs influence communities over extended periods.
+
+In conclusion, while the PollinateTO dataset provides an important resource for examining urban gardening initiatives, careful consideration of the sampling and observational data structure is essential when interpreting the findings.
+The potential biases—such as selection bias, omitted variable bias, and confounding bias—must be accounted for when drawing conclusions about the factors influencing garden size.
+Future research could benefit from additional data sources, such as surveys of garden participants or longitudinal data, to further explore these relationships and improve the robustness of the findings.
 
@@
 The posterior predictive check (pp_check) helps assess how well the model fits the observed data in the context of your research on the factors influencing community garden funding by PollinateTO across different Toronto neighborhoods.
-By comparing the model’s predictions (posterior distribution) with the observed data, this diagnostic reveals whether the model is accurately capturing the key influences, such as garden type, funding year, and neighborhood characteristics.
+By comparing the model’s predictions (posterior distribution) with the observed data, this diagnostic shows whether the model is accurately capturing the key influences, such as garden type, funding year, and neighborhood characteristics.
 A good fit suggests that these factors are well-represented in the model, while significant deviations between the predictions and the data may indicate that the model needs refinement.
@@
 A well-calibrated model should show a substantial difference between the prior and posterior for key parameters, indicating that the observed data has influenced the model’s estimates.
-This comparison is crucial for understanding which variables are most influential in determining garden size and funding across neighborhoods.
+This comparison is important for understanding which variables are most influential in determining garden size and funding across neighborhoods.
 
@@
 Non-convergence, shown by trends or failure to stabilize, may result from poor model specification or insufficient warmup.
-Good mixing is characterized by rapid oscillations and variability, reflecting effective exploration of the parameter space, whereas slow mixing or long stretches of similar values may indicate inefficiency or high autocorrelation.
-Proper convergence and mixing are critical to draw robust inferences about how factors like indigenous-led status or funding year influence garden size, ensuring the validity of your findings.
+Good mixing is characterized by rapid oscillations and variability, reflecting effective understanding of the parameter space, whereas slow mixing or long stretches of similar values may indicate inefficiency or high autocorrelation.
+Proper convergence and mixing are important to draw robust inferences about how factors like indigenous-led status or funding year influence garden size, ensuring the validity of your findings.
 This suggests...
@@
 @fig-rhatplot is a Rhat plot.
-Rhat plots are critical diagnostics for assessing convergence in Bayesian logistic regression models like this one.
+Rhat plots are important diagnostics for assessing convergence in Bayesian logistic regression models like this one.
 The Rhat statistic, or Gelman-Rubin diagnostic, compares within-chain and between-chain variance to determine if the Markov chains have converged to the posterior distribution.
@@
 By ensuring all parameters in the model have Rhat values near 1, the plot provides confidence in the stability and reliability of the posterior estimates.
-This is crucial for interpreting how predictors like indigenous-led status and funding year influence garden size accurately.
+This is important for interpreting how predictors like indigenous-led status and funding year influence garden size accurately.
 
